---
title: "BiodiversityModelPrediction_Fish_Software_Release"
author: "Jenny Rogers"
date: "2024-02-05"
output: html_document
---


```{r}

library(tidyverse)
library(sf)
library(glmmTMB)
library(readxl)



```

proportional abundance and probabilty of zero predictions
```{r}

load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates.RData")

#need to scale the variables
test <- NHDplusV2_NewEngCrop %>% 
  data.frame() %>% 
  select(-geometry)

test <- test[,c(7:8, 10:42)]

#we need to scale these variables, but first I will calculate the mean and sd of each column, bc we will need to use these values to scale these variables in the predicted scenarios

means <- sapply(test, mean, na.rm = T)
save(means, file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_mean.RData")
sd <- sapply(test, sd, na.rm = T)
save(sd, file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_sd.RData")

#scale data
test <- data.frame(scale(test))

NHDplusV2_NewEngCrop <- cbind((NHDplusV2_NewEngCrop)[,c(1:6, 9)], test)



#watershed shape files
huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  select(name, tnmid)


#fish traits
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fish_traits.RData")
temperature <- read_xlsx("C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/thermalpref_origin_tolerance.xlsx",
                  sheet = 1,
                  range = cell_cols("A:G")) %>% 
  pivot_longer(2:7, names_to = "state", values_to = "tmp") %>% 
  filter(!is.na(tmp),
         !tmp == "eury",
         !(common_name %in% c("american brook lamprey") & tmp == "cold"),
         !(common_name %in% c("common shiner", "white sucker", "smallmouth bass", "fallfish") & tmp == "warm")) %>% 
  select(-state) %>% 
  unique()

origin <- read_xlsx("C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/thermalpref_origin_tolerance.xlsx",
                  sheet = 2,
                  range = cell_cols("A:G"))%>% 
  pivot_longer(2:7, names_to = "state", values_to = "origin") %>% 
  filter(!is.na(origin),
         !(common_name == "spottail shiner" & origin == "introduced"),
         !origin == "unk") %>% 
  select(-state) %>%  
  unique()

#load fish clusters
geogclust <- read.csv("tmpfigures/fishclusters_bygeography.csv")
geogclust <- geogclust %>% 
  select(-1) %>% 
  rename(geogclust = final.cluster)
climclust <- read.csv("tmpfigures/fishclusters_byclimate.csv")
climclust <- climclust %>% 
  select(-1) %>% 
  rename(climclust = final.cluster)
tempclust <- read.csv("tmpfigures/fishclusters_bytemperature.csv")
tempclust <- tempclust %>% 
  select(-1) %>% 
  rename(tempclust = final.cluster)
flowclust <- read.csv("tmpfigures/fishclusters_bystreamflow.csv")
flowclust <- flowclust %>% 
  select(-1) %>% 
  rename(flowclust = final.cluster)

#proportional abundance predictions

file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "^propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop, type = "response"))
names(predictions)[1] <- file.list[i]


for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
    tmp <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  

  
}


#join the predictions to the list of comids
NHDplusV2_NewEngCrop_baseline <- cbind(NHDplusV2_NewEngCrop, predictions)


ggplot(data = NHDplusV2_NewEngCrop_baseline)+
  geom_sf(aes(color = propabun_fallfish.rds))+
  scale_color_gradient2(low = "blue",
                        mid = "yellow",
                        high = "red",
                        midpoint = 0.5)



#align prediction results with the unscaled covariates and make plots
load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates.RData")
pred <- cbind(NHDplusV2_NewEngCrop, predictions)
pred <- pred %>%
  data.frame %>% 
  pivot_longer(43:95, values_to = "propabun", names_to = "common_name") 
pred$common_name <- gsub("propabun_", "", pred$common_name)
pred$common_name <- gsub(".rds", "", pred$common_name)
pred$common_name <- gsub("\\.", " ", pred$common_name)


#join to trait data and clusters so we can filter for certain temperature groups or origin
pred <- pred %>% 
  left_join(temperature, by = "common_name") %>% 
  left_join(origin, by = "common_name") %>% 
  left_join(geogclust, by = "common_name") %>% 
  left_join(climclust, by = "common_name") %>% 
  left_join(final_traits, by = "common_name") %>% 
  left_join(flowclust, by = "common_name") %>% 
  left_join(tempclust, by = "common_name")

#write csv with final spp groupings
sppgroups <- pred %>% 
  select(common_name, tmp, tolerance, tempclust, flowclust, origin) %>% 
  unique() %>% 
  arrange(tmp, tolerance, tempclust, flowclust, origin)
sppgroups$origin[sppgroups$origin == "introduced"] <- "I"
sppgroups$origin[sppgroups$origin == "native"] <- "N"
sppgroups$tolerance[sppgroups$tolerance == "fluvial specialist"] <- "FS"
sppgroups$tolerance[sppgroups$tolerance == "fluvial dependent"] <- "FD"
sppgroups$tolerance[sppgroups$tolerance == "generalist"] <- "MG"
#add in scienfific names
names <- read_excel("C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/fishspp_lookup_rq_comments.xlsx") %>% 
  select(scientific_name, common_name) 
sppgroups <- left_join(sppgroups, names, by = "common_name")%>% 
  select(common_name, scientific_name, tmp, tolerance, tempclust, flowclust, origin)
sppgroups$common_name <- str_to_title(sppgroups$common_name)
sppgroups$scientific_name <- str_to_sentence(sppgroups$scientific_name)
write.csv(sppgroups, "tmpfigures/paper1/sppgroups.csv")




#ratio of native to non-native species
OR <- pred %>% 
  select(COMID, huc12_name, huc12_tnmid, common_name, origin, propabun) %>% 
  unique() %>% 
  filter(!is.na(origin),
         propabun> 0.01) %>% 
  group_by(COMID, huc12_name, huc12_tnmid, origin) %>% 
  summarise(propabun = sum(propabun, na.rm = T)) %>% 
  ungroup() %>% 
  pivot_wider(values_from = propabun, names_from = origin) %>% 
  mutate(NatIntRatio = native/introduced) %>% 
  group_by(huc12_name, huc12_tnmid) %>%
  summarise(NatIntRatio = mean(NatIntRatio, na.rm = T))
OR <- left_join(huc12, OR, by = c("name" = "huc12_name", "tnmid" = "huc12_tnmid")) %>% 
  filter(!is.na(NatIntRatio))
 
aa <- ggplot(data = OR)+
  geom_sf(aes(fill = NatIntRatio), color = NA)+
  scale_fill_gradient2(low = "yellow",
                        mid = "orange",
                        high = "blue",
                        midpoint = 50,
                       limits = c(1,100), 
                       breaks = c(1,100))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        legend.text = element_text(size = 5),
        legend.title = element_text(size = 5),
        legend.position = c(0.78, 0.21),
        legend.key.size = unit(.25, "cm"),
        title = element_text(size = 10),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.margin = unit(c(1, 1, 1, 1), "mm"))+
  ggtitle("D) Native to Introduced")+
  labs(fill = "Ratio")

ggsave(aa, file = "tmpfigures/paper1/NatIntRatio.png", 
    width = 18,
    height = 15,
    units = "cm",
    dpi = 300)


#ratio of specialist to generalist species
SP <- pred %>% 
  select(COMID, huc12_name, huc12_tnmid, common_name, tolerance, propabun) %>% 
  unique() %>% 
  filter(!is.na(tolerance),
         propabun> 0.001,
         tolerance %in% c("fluvial specialist", "generalist")) %>% 
  group_by(COMID, huc12_name, huc12_tnmid, tolerance) %>% 
  summarise(propabun = sum(propabun, na.rm = T)) %>% 
  ungroup() %>% 
  pivot_wider(values_from = propabun, names_from = tolerance) %>% 
  mutate(speGenRatio = `fluvial specialist`/generalist) %>% 
  group_by(huc12_name, huc12_tnmid) %>%
  summarise(speGenRatio = mean(speGenRatio, na.rm = T))
SP <- left_join(huc12, SP, by = c("name" = "huc12_name", "tnmid" = "huc12_tnmid")) %>% 
  filter(!is.na(speGenRatio))
 
bb <- ggplot(data = SP)+
  geom_sf(aes(fill = speGenRatio), color = NA)+
  scale_fill_gradient2(low = "yellow",
                        mid = "orange",
                        high = "blue",
                        midpoint = 330,
                       limits = c(0,665), 
                       breaks = c(0,665))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        title = element_text(size = 10),
        axis.text = element_blank(),
        axis.ticks = element_blank())+
  ggtitle("B) Specialist to Generalist")+
  labs(fill = "Ratio")

ggsave(bb, file = "tmpfigures/paper1/SpeGenRatio.png", 
    width = 18,
    height = 15,
    units = "cm",
    dpi = 300)

fig4 <- gridExtra::grid.arrange(aa,bb, nrow = 1, ncol = 2)
ggsave(fig4, file = "tmpfigures/paper1/fig4.png")




#Temperature Clusters
TC <- pred %>% 
  select(COMID, huc12_name, huc12_tnmid, common_name, tempclust, propabun) %>% 
  unique() %>% 
  filter(!is.na(tempclust),
         common_name != "brook trout") %>% 
  group_by(COMID, huc12_name, huc12_tnmid, tempclust) %>% 
  summarise(propabun = sum(propabun, na.rm = T)) %>% 
  ungroup() %>% 
  group_by(huc12_name, huc12_tnmid, tempclust) %>%
  summarise(propabun = mean(propabun, na.rm = T))
TC <- left_join(huc12, TC, by = c("name" = "huc12_name", "tnmid" = "huc12_tnmid")) %>% 
  filter(!is.na(tempclust))
TC$propabun <- ifelse(TC$propabun > 1, 1, TC$propabun)  
TC$tempclust[TC$tempclust == "1"] <- "1-cold, mid elev"
TC$tempclust[TC$tempclust == "2"] <- "2-warm, sm sheds"
TC$tempclust[TC$tempclust == "3"] <- "3-cool, low elev"
TC$tempclust[TC$tempclust == "4"] <- "4-cool, mid elev"
TC$tempclust[TC$tempclust == "5"] <- "5-cold, high elev"
TC$tempclust[TC$tempclust == "6"] <- "6-warm, lg sheds"

a <- ggplot(data = TC)+
  geom_sf(aes(fill = propabun), color = NA)+
  scale_fill_gradient2(low = "yellow",
                        mid = "orange",
                        high = "blue",
                        midpoint = .5,
                       limits = c(0,1), 
                       breaks = c(0,1))+
  facet_wrap(~tempclust)+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.position = c(0.94, 0.10),
        legend.key.size = unit(.2, "cm"))+
  ggtitle("A) Temperature Clusters")+
  labs(fill = "P.Abun.")

ggsave(a, file = "tmpfigures/paper1/TempClusters.png", 
    width = 18,
    height = 15,
    units = "cm",
    dpi = 300)

#streamflow Clusters
FC <- pred %>% 
  select(COMID, huc12_name, huc12_tnmid, common_name, flowclust, propabun) %>% 
  unique() %>% 
  filter(!is.na(flowclust),
         common_name != "brook trout") %>% 
  group_by(COMID, huc12_name, huc12_tnmid, flowclust) %>% 
  summarise(propabun = sum(propabun, na.rm = T)) %>% 
  ungroup() %>% 
  group_by(huc12_name, huc12_tnmid, flowclust) %>%
  summarise(propabun = mean(propabun, na.rm = T))
FC <- left_join(huc12, FC, by = c("name" = "huc12_name", "tnmid" = "huc12_tnmid")) %>% 
  filter(!is.na(flowclust))
FC$propabun <- ifelse(FC$propabun > 1, 1, FC$propabun)
FC$flowclust[FC$flowclust == "1"] <- "1-Hi flood, Med Q"
FC$flowclust[FC$flowclust == "2"] <- "2-Mod flood, Lo Q"
FC$flowclust[FC$flowclust == "3"] <- "3-Hi flood, Low Q"
FC$flowclust[FC$flowclust == "4"] <- "4-Mod flood, Med Q"
FC$flowclust[FC$flowclust == "5"] <- "5-Hi flood, Hi Q"
FC$flowclust[FC$flowclust == "6"] <- "6-Lo flood, Med Q"

b <- ggplot(data = FC)+
  geom_sf(aes(fill = propabun), color = NA)+
  scale_fill_gradient2(low = "yellow",
                        mid = "orange",
                        high = "blue",
                        midpoint = .5,
                       limits = c(0,1), 
                       breaks = c(0,1))+
  facet_wrap(~flowclust, nrow = 2)+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.position = c(0.94, 0.10),
        legend.key.size = unit(.2, "cm"),
        plot.margin = unit(c(1, 1, 1, 1), "mm"))+
  ggtitle("B) Streamflow Clusters")+
  labs(fill = "P.Abun.")

ggsave(b, file = "tmpfigures/paper1/FlowClusters.png", 
    width = 18,
    height = 15,
    units = "cm",
    dpi = 300)


fig7 <- gridExtra::grid.arrange(a,b, nrow = 1, ncol = 2)
ggsave(fig7, file = "tmpfigures/paper1/fig7.png")


#Temperature Guilds
TA <- pred %>% 
  select(COMID, huc12_name, huc12_tnmid, common_name, tmp, propabun) %>% 
  unique() %>% 
  filter(!is.na(tmp),
         common_name != "brook trout") %>% 
  group_by(COMID, huc12_name, huc12_tnmid, tmp) %>% 
  summarise(propabun = sum(propabun, na.rm = T)) %>% 
  ungroup() %>% 
  group_by(huc12_name, huc12_tnmid, tmp) %>%
  summarise(propabun = mean(propabun, na.rm = T))
TA <- left_join(huc12, TA, by = c("name" = "huc12_name", "tnmid" = "huc12_tnmid")) %>% 
  filter(!is.na(tmp))
TA$propabun <- ifelse(TA$propabun > 1, 1, TA$propabun) 
TA$tmp <- str_to_title(TA$tmp)


c <- ggplot(data = TA)+
  geom_sf(aes(fill = propabun), color = NA)+
  scale_fill_gradient2(low = "yellow",
                        mid = "orange",
                        high = "blue",
                        midpoint = .5,
                       limits = c(0,1), 
                       breaks = c(0,1))+
  facet_wrap(~tmp)+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        legend.text = element_text(size = 5),
        strip.text = element_text(size = 8),
        title = element_text(size = 10),
        legend.position = c(0.93, 0.25),
        legend.key.size = unit(.25, "cm"),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.margin = unit(c(1, 1, 1, 1), "mm"))+
  ggtitle("A) Temperature Guilds")+
  labs(fill = NULL)

ggsave(c, file = "tmpfigures/paper1/TempGuilds.png", 
    width = 18,
    height = 10,
    units = "cm",
    dpi = 300)

#Fluvial guilds
pred$tolerance[pred$tolerance == "generalist"] <- "Macrohab. Generalist"
FA <- pred %>% 
  select(COMID, huc12_name, huc12_tnmid, common_name, tolerance, propabun) %>% 
  unique() %>% 
  filter(!is.na(tolerance),
         common_name != "brook trout") %>% 
  group_by(COMID, huc12_name, huc12_tnmid, tolerance) %>% 
  summarise(propabun = sum(propabun, na.rm = T)) %>% 
  ungroup() %>% 
  group_by(huc12_name, huc12_tnmid, tolerance) %>%
  summarise(propabun = mean(propabun, na.rm = T))
FA <- left_join(huc12, FA, by = c("name" = "huc12_name", "tnmid" = "huc12_tnmid")) %>% 
  filter(!is.na(tolerance))
FA$propabun <- ifelse(FA$propabun > 1, 1, FA$propabun) 
FA$tolerance <- str_to_title(FA$tolerance)


d <- ggplot(data = FA)+
  geom_sf(aes(fill = propabun), color = NA)+
  scale_fill_gradient2(low = "yellow",
                        mid = "orange",
                        high = "blue",
                        midpoint = .5,
                       limits = c(0,1), 
                       breaks = c(0,1))+
  facet_wrap(~tolerance)+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        legend.text = element_text(size = 5),
        strip.text = element_text(size = 8),
        legend.position = c(0.93, 0.25),
        legend.key.size = unit(.25, "cm"),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(size = 10),
        plot.margin = unit(c(1, 1, 1, 1), "mm"))+
  ggtitle("B) Habitat Guilds")+
  labs(fill = NULL)

ggsave(d, file = "tmpfigures/paper1/FlowGuilds.png", 
    width = 18,
    height = 10,
    units = "cm",
    dpi = 300)

#Origin groups
OG <- pred %>% 
  select(COMID, huc12_name, huc12_tnmid, common_name, origin, propabun) %>% 
  unique() %>% 
  filter(!is.na(origin),
         common_name != "brook trout") %>% 
  group_by(COMID, huc12_name, huc12_tnmid, origin) %>% 
  summarise(propabun = sum(propabun, na.rm = T)) %>% 
  ungroup() %>% 
  group_by(huc12_name, huc12_tnmid, origin) %>%
  summarise(propabun = mean(propabun, na.rm = T))
OG <- left_join(huc12, OG, by = c("name" = "huc12_name", "tnmid" = "huc12_tnmid")) %>% 
  filter(!is.na(origin))
OG$propabun <- ifelse(OG$propabun > 1, 1, OG$propabun) 
OG$origin <- str_to_title(OG$origin)



OG2 <- ggplot(data = OG)+
  geom_sf(aes(fill = propabun), color = NA)+
  scale_fill_gradient2(low = "yellow",
                        mid = "orange",
                        high = "blue",
                        midpoint = .5,
                       limits = c(0,1), 
                       breaks = c(0,1))+
  facet_wrap(~origin)+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        legend.text = element_text(size = 5),
        strip.text = element_text(size = 8),
        legend.position = c(0.90, 0.25),
        legend.key.size = unit(.25, "cm"),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(size = 10),
        plot.margin = unit(c(1, 1, 1, 1), "mm"))+
  ggtitle("D) Origin Groups")+
  labs(fill = NULL)



#Brook Trout
BT <- pred %>% 
  select(COMID, huc12_name, huc12_tnmid, common_name, propabun) %>% 
  unique() %>% 
  filter(common_name == "brook trout") %>%
  group_by(huc12_name, huc12_tnmid) %>%
  summarise(propabun = mean(propabun, na.rm = T))
BT <- left_join(huc12, BT, by = c("name" = "huc12_name", "tnmid" = "huc12_tnmid"))
BT$propabun <- ifelse(BT$propabun > 1, 1, BT$propabun)  


e <- ggplot(data = BT)+
  geom_sf(aes(fill = propabun), color = NA)+
  scale_fill_gradient2(low = "yellow",
                        mid = "orange",
                        high = "blue",
                        midpoint = .5,
                       limits = c(0,1), 
                       breaks = c(0,1))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        title = element_text(size = 10),
        plot.margin = unit(c(1, 1, 1, 1), "mm"))+
  ggtitle("C) Brook Trout")+
  labs(fill = "Prop. Abun.")
  

ggsave(e, file = "tmpfigures/paper1/BrookTrout.png", 
    width = 8,
    height = 7,
    units = "cm",
    dpi = 300)

#add a locational map of new england in the uS for figure 2 in paper 1
#watershed shape files
huc8 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU8/WBDHU8_NE.shp")
#load new england state data
states <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/newenglandshape/NEWENGLAND_POLY.shp")
#load USA states
US <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/USstates/cb_2018_us_state_5m.shp")
US <- US %>% 
  filter(STUSPS %in% c("ME", "NH", "VT", "MA", "CT", "RI", "NY", "PA", "NJ", "MD", "VA", "OH", "WV", "DE"))
st_crs(huc8) == st_crs(US)
states <- st_transform(states, crs = st_crs(huc8))
st_crs(huc8) == st_crs(states)


locationalMap <- ggplot(data = US)+
  geom_sf(color = "black", fill = "grey90")+
  geom_sf(data = states, fill = "red2")+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        axis.text = element_text(size = 7),
        legend.position = "none",
        title = element_text(size = 10))+
  ggtitle("E) Study Location")


fig2 <- gridExtra::grid.arrange(arrangeGrob(c,d, ncol = 2),
                                arrangeGrob(e, OG2, locationalMap, ncol = 3))
ggsave(fig2, file = "tmpfigures/paper1/fig2.png", width = 18, height = 12, units = "cm")


#plot each variable calcuated above against all the others in the ggpairs plot
TC <- TC %>% 
  data.frame() %>% 
  select(name, tnmid, tempclust, propabun)  %>%
  unique() %>% 
  pivot_wider(names_from = tempclust, values_from = propabun) %>% 
  rename(Tempclust_1 = "1-cold, mid elev", Tempclust_2 = "2-warm, sm sheds", Tempclust_3 = "3-cool, low elev", 
         Tempclust_4 = "4-cool, mid elev", Tempclust_5 = "5-cold, high elev", Tempclust_6 = "6-warm, lg sheds")

FC <- FC %>% 
  data.frame() %>% 
  select(name, tnmid, flowclust, propabun)  %>%
  unique() %>% 
  pivot_wider(names_from = flowclust, values_from = propabun) %>% 
  rename(Flowclust_1 = "1-Hi flood, Med Q", Flowclust_2 = "2-Mod flood, Lo Q", Flowclust_3 = "3-Hi flood, Low Q", 
         Flowclust_4 = "4-Mod flood, Med Q", Flowclust_5 = "5-Hi flood, Hi Q", Flowclust_6 = "6-Lo flood, Med Q")

TA <- TA %>% 
  data.frame() %>% 
  select(name, tnmid, tmp, propabun)  %>%
  unique() %>% 
  pivot_wider(names_from = tmp, values_from = propabun) 

FA <- FA %>% 
  data.frame() %>% 
  select(name, tnmid, tolerance, propabun)  %>%
  unique() %>% 
  pivot_wider(names_from = tolerance, values_from = propabun)

BT <- BT %>% 
  data.frame() %>%
  select(name, tnmid, propabun)  %>%
  unique() %>% 
  filter(!is.na(propabun)) %>% 
  rename("Brook Trout" = propabun)

OR <- OR %>% 
  data.frame() %>% 
  select(name, tnmid, NatIntRatio) %>% 
  unique() %>% 
  filter(!is.na(NatIntRatio))

OG <- OG %>% 
  data.frame() %>% 
  select(name, tnmid, origin, propabun) %>% 
  unique() %>% 
  pivot_wider(names_from = origin, values_from = propabun)

SP <- SP %>% 
  data.frame() %>% 
  select(name, tnmid, speGenRatio) %>% 
  unique() %>% 
  filter(!is.na(speGenRatio))

compare <- left_join(TC, FC, by = c("name", "tnmid")) %>% 
  left_join(TA, by = c("name", "tnmid")) %>% 
  left_join(FA, by = c("name", "tnmid")) %>% 
  left_join(BT, by = c("name", "tnmid")) %>%
  left_join(OG, by = c("name", "tnmid")) %>%  
  select(-name, -tnmid)
  
compare <- compare[complete.cases(compare),]

library(GGally)
ggpairs(compare)
names(compare)

names(compare)[names(compare) == "Macrohab. Generalist"] <-  "Macrohabitat Generalist"
test <- cor(compare)
test <- test %>% 
  data.frame() %>% 
  select(13:21) %>% 
  filter(rownames(test) %in% c("Tempclust_1", "Tempclust_2", "Tempclust_3", "Tempclust_4", 
                               "Tempclust_5", "Tempclust_6", "Flowclust_1", "Flowclust_2", 
                               "Flowclust_3", "Flowclust_4", "Flowclust_5", "Flowclust_6")) %>%
  rename("Fluvial Specialist"="Fluvial.Specialist",  "Fluvial Dependent" = "Fluvial.Dependent", "Brook Trout" = "Brook.Trout",
         "Macrohabitat Generalist" = "Macrohabitat.Generalist") %>% 
  as.matrix()

library(corrplot)
png(file = "tmpfigures/paper1/correlationmatrix.png", res = 200, width = 10, height = 15, units = "in", pointsize = 24)
corrplot(test, method = "square", outline = T, addgrid.col = "darkgray", 
         addCoef.col = "white", number.digits = 2, number.cex = 0.9, tl.col = "black")
dev.off()

write.csv(test, file = "tmpfigures/paper1/correlationmatrix.csv")






#make violin plots with the distrubiton of each trait by lat and long
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDv2_huc_state_join.RData")

#join the predictions_zero to the list of comids
NHDplusV2_NewEngCrop <- NHDplusV2_NewEngCrop %>% 
  select(COMID)
bylat <- cbind(NHDplusV2_NewEngCrop, predictions) 
bylat <- data.frame(bylat) %>% 
  select(-geometry) %>% 
  pivot_longer(2:54, values_to = "propabun", names_to = "common_name") 
bylat$common_name <- gsub("propabun_", "", pred$common_name)
bylat$common_name <- gsub(".rds", "", pred$common_name)
bylat$common_name <- gsub("\\.", " ", pred$common_name)


#join to trait data and clusters so we can filter for certain temperature groups or origin
bylat <- bylat %>% 
  left_join(temperature, by = "common_name") %>% 
  left_join(origin, by = "common_name") %>% 
  left_join(geogclust, by = "common_name") %>% 
  left_join(climclust, by = "common_name") %>% 
  left_join(final_traits, by = "common_name")%>% 
  left_join(flowclust, by = "common_name") %>% 
  left_join(tempclust, by = "common_name")

bylat <- left_join(NHDv2_huc_state_join, bylat, by = "COMID")

df1 <- bylat %>% 
  data.frame() %>% 
  group_by(COMID, tmp) %>% 
  summarize(propabun = sum(propabun, na.rm = T)) %>% 
  rename(trait = tmp)
df2 <- bylat %>% 
  data.frame() %>% 
  group_by(COMID, tolerance) %>% 
  summarize(propabun = sum(propabun, na.rm = T)) %>% 
  rename(trait = tolerance)
df3 <- bylat %>% 
  data.frame() %>% 
  group_by(COMID, tempclust) %>% 
  summarize(propabun = sum(propabun, na.rm = T)) %>% 
  rename(trait = tempclust) %>% 
  mutate(trait = as.character(trait) ) %>% 
  filter(!is.na(trait))
df3$trait <- paste("TC", df3$trait, sep = "_")
df4 <- bylat %>% 
  data.frame() %>% 
  group_by(COMID, flowclust) %>% 
  summarize(propabun = sum(propabun, na.rm = T)) %>% 
  rename(trait = flowclust) %>% 
  mutate(trait = as.character(trait) )%>% 
  filter(!is.na(trait))
df4$trait <- paste("FC", df4$trait, sep = "_")

df5 <- bylat %>% 
  data.frame() %>% 
  filter(common_name == "brook trout") %>% 
  select(COMID, common_name, propabun) %>% 
  rename(trait = common_name)
df6 <- bylat %>% 
  data.frame() %>% 
  select(COMID, common_name, origin, propabun) %>% 
  group_by(COMID, origin) %>% 
  summarise(propabun = sum(propabun, na.rm = T)) %>% 
  rename(trait = origin) %>% 
  filter(!is.na(trait))


latlong <- NHDv2_huc_state_join %>%
  data.frame() %>% 
  select(COMID, lat, long) %>% 
  unique()
load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates.RData")
elev <- NHDplusV2_NewEngCrop %>% 
  select(COMID, ElevCat)

df7 <- rbind(df1, df2, df3, df4, df5, df6)
df7 <- left_join(df7, latlong, by = "COMID") %>% 
  left_join(elev, by = "COMID") %>% 
  filter(!is.na(trait),
         !is.na(ElevCat)) %>% 
  mutate(latitude  = ifelse(lat > 44.51944, "High latitude",
                            ifelse(lat < 42.97004, "Low latitude", "Mid latitude"))) %>% 
  mutate(elevation  = ifelse(ElevCat > 181.4322 , "High elevation", "Low elevation")) %>% 
  mutate(longitude  = ifelse(long > -71.27307, "East", "West")) %>% 
  filter(propabun < 1.2)

df7$latitude <- factor(df7$latitude, levels = c("Low latitude", "Mid latitude", "High latitude"))
df7$elevation <- factor(df7$elevation, levels = c("Low elevation", "Mid elevation", "High elevation"))

df8 <- df7 %>% 
  filter(trait %in% c("TC_1", "TC_2", "TC_3", "TC_4", "TC_5", "TC_6", 
                      "FC_1", "FC_2", "FC_3", "FC_4", "FC_5", "FC_6"))

df9 <- df7 %>% 
  filter(trait %in% c("cold", "cool", "warm", "fluvial specialist", "fluvial dependent", "generalist", 
                      "brook trout","native", "introduced" ))
df9$trait[df9$trait == "native"] <- "Native"
df9$trait[df9$trait == "introduced"] <- "Introduced"
df9$trait[df9$trait == "generalist"] <- "Macrohabitat Generalist"

ggplot(data = df8, mapping = aes(x = trait, y = propabun))+
  geom_boxplot(outlier.size = .5, outlier.colour = "grey80")+ 
  facet_grid(rows = vars(elevation), cols = vars(latitude))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(x = "Grouping", y = "Average Proportional Abundance")+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12))+
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red")+
  geom_hline(yintercept = 0.25, linetype = "dashed", color = "red") +
  geom_hline(yintercept = 0.75, linetype = "dashed", color = "red")
ggsave(plot = last_plot(),
       file = "tmpfigures/paper1/clustersbylatlong.png", 
    width = 20,
    height = 12,
    units = "cm",
    dpi = 300)

df9$trait <- str_to_title(df9$trait)
df9$trait <- factor(df9$trait, levels = c("Cold", "Cool", "Warm", 
                                          "Fluvial Specialist", "Fluvial Dependent","Macrohabitat Generalist",
                                          "Brook Trout", "Native", "Introduced" ))

ggplot(data = df9, mapping = aes(x = trait, y = propabun))+
  geom_boxplot(outlier.size = .5, outlier.colour = "grey80")+
  facet_grid(rows = vars(elevation), cols = vars(latitude))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(x = "Grouping", y = "Average Proportional Abundance")+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12))+
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  geom_hline(yintercept = 0.25, linetype = "dashed", color = "red") +
  geom_hline(yintercept = 0.75, linetype = "dashed", color = "red")
ggsave(plot = last_plot(),
       file = "tmpfigures/paper1/Metricbylatlong.png", 
    width = 20,
    height = 12,
    units = "cm",
    dpi = 300)


 