---
title: "BiodiversityModelPrediction_Mussel"
author: "Jenny Rogers"
date: "2023-04-14"
output: html_document
---

```{r}
#mussel data
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/mussel_occurrence.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/mussel_count_with_zeros.RData")
mussel_shp <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/sppdata/all_mussel_event.shp")

#mussel event and hydrography join
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/mussel_event_huc_join.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/mussel_event_flowline_join.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/mussel_event_flowlineV2_join.RData")

#watershed shape files
huc8 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU8/WBDHU8_NE.shp")
huc10 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU10/WBDHU10_NE.shp")

#load New England state data
states <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/newenglandshape/NEWENGLAND_POLY.shp")


```

Baseline predictions


need to calculate the baseline host fish distributions at the HUC10 scale (we also did this in the biodiveristyModel_mussel file)

Host fish distributions at HUC10 scale.  This predicts the host fish distributions in the baseline condition. 
When we want to do mussel predictions in future time periods, we will need to redo this calcuation changing out the covariate data file for one with climate change and/or the management scenarios.  The new data sets need to be scaled according to the same mean and sd as the baseline.
```{r}

load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates.RData")

#need to scal the variavles
test <- NHDplusV2_NewEngCrop %>% 
  data.frame() %>% 
  select(-geometry)

test <- test[,c(7:8, 10:39)]

#scale data
test <- data.frame(scale(test))

NHDplusV2_NewEngCrop <- cbind((NHDplusV2_NewEngCrop)[,c(1:6, 9)], test)


#fish host covariate with probability of presence by spp

file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "pres_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop, type = "response"))
names(predictions)[1] <- file.list[i]




for (i in 2:57) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  
  
}


#join the predictions to the list of comids
NHDplusV2_NewEngCrop_baseline <- cbind(NHDplusV2_NewEngCrop, predictions)

#average by HUC10 becuase that is the scale we are modeling mussels
fish_host_prob_occ <- NHDplusV2_NewEngCrop_baseline %>%
  data.frame() %>% 
  select(huc10_name, huc10_tnmid, 40:96) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc10_name)) %>%
  pivot_longer(3:59, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc10_name, huc10_tnmid) %>% 
  summarise(total_prob_occ = sum(prob_occ))
  





#fish host covariate with proportional abundance by family. First do prop abund model. then add within a family


file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop, type = "response"))
names(predictions)[1] <- file.list[i]




for (i in 2:57) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  
  
}


#join the predictions to the list of comids
NHDplusV2_NewEngCrop_baseline <- cbind(NHDplusV2_NewEngCrop, predictions)

#average by HUC10 becuase that is the scale we are modeling mussels
fish_host_propabun <- NHDplusV2_NewEngCrop_baseline %>%
  data.frame() %>% 
  select(huc10_name, huc10_tnmid, 40:96) %>% 
  filter(!is.na(propabun_brook.trout.rds),
         !is.na(huc10_name)) %>% 
  group_by(huc10_name, huc10_tnmid) %>% 
  summarise_all(mean) %>% 
  pivot_longer(3:59, names_to = "common_name", values_to = "propabun")
fish_host_propabun$common_name <- gsub("propabun_", "", fish_host_propabun$common_name)
fish_host_propabun$common_name <- gsub(".rds", "", fish_host_propabun$common_name)
fish_host_propabun$common_name <- gsub("\\.", " ", fish_host_propabun$common_name)

#add family information
fish_host_propabun <- fish_host_propabun %>% 
  mutate(family = ifelse(common_name%in% c("american brook lamprey"), "Petromyzontidae",
                         ifelse(common_name %in% c("american eel"), "Anguillidae",
                                ifelse(common_name%in% c("atlantic salmon", "brook trout", "brown trout", "rainbow trout", "round whitefish"), "Salmonidae",
                                       ifelse(common_name %in% c("banded killifish"), "Fundulidae",
                                              ifelse(common_name %in% c("banded sunfish", "black crappie", "bluegill", "green sunfish", "largemouth bass", "pumpkinseed", "redbreast sunfish", "rock bass", "smallmouth bass"), "Centrarchidae",
                                                     ifelse(common_name %in% c("bluntnose minnow", "bridle shiner", "common carp", "common shiner", "creek chub", "cutlips minnow", "eastern blacknose dace", "eastern silvery minnow", "fallfish", "fathead minnow", "finescale dace", "golden shiner", "lake chub", "longnose dace", "mimic shiner", "redbelly dace", "rosyface shiner", "rosyside dace", "spotfin shiner", "spottail shiner"), "Cyprinidae",
                                                            ifelse(common_name %in% c("bowfin"), "Amiidae",
                                                                   ifelse(common_name %in% c("brook stickleback"), "Gasterosteidae",
                                                                          ifelse(common_name %in% c("brown bullhead", "channel catfish", "margined madtom", "white catfish"), "Ictaluridae",
                                                                                 ifelse(common_name %in% c("burbot"), "Lotidae",
                                                                                        ifelse(common_name %in% c("central mudminnow"), "Umbridae" ,
                                                                                               ifelse(common_name %in% c("chain pickerel", "northern pike", "redfin pickerel"), "Esocidae",
                                                                                                      ifelse(common_name %in% c("creek chubsucker", "longnose sucker", "white sucker", "yellow bullhead"), "Catostomidae",
                                                                                                             ifelse(common_name %in% c("slimy sculpin"), "Cottidae", "Percidae")
                                                                                                             ))))))))))))))

#swamp darter", "tessellated darter", "walleye", "yellow perch" are percidae

#group by family for total proportional abundance by family
fish_host_propabun <- fish_host_propabun %>% 
  group_by(huc10_name, huc10_tnmid, family) %>% 
  summarise(propabun = sum(propabun)) %>% 
  pivot_wider(names_from = family, values_from = propabun) %>% 
  select(-Gasterosteidae) #its zero in every observation





```




Load in the HUC10 mussel covariates,  join the host fish distributions, and scale
```{r}

load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_mussel_covariates_HUC10.RData")

covariates <- left_join(NHDplusV2_NewEngCrop_mussel_covariates_HUC10, fish_host_prob_occ, by = c("huc10_name", "huc10_tnmid"))

covariates <- left_join(covariates, fish_host_propabun, by = c("huc10_name", "huc10_tnmid"))


#scale mussel covariates
test <- covariates[,c(3:45)]

means <- sapply(test, mean, na.rm = T)
save(means, file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_mean.RData")
sd <- sapply(test, sd, na.rm = T)
save(sd, file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_sd.RData")


#scale data
test <- data.frame(scale(test))

covariates <- cbind((covariates)[,c(1:2)], test)





```

Load in the mussel models and apply them to the baseline data

```{r}

file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "mussel_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = predict(mdl, newdata = covariates, type = "response"))
names(predictions)[1] <- file.list[i]




for (i in 2:12) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = covariates, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  
  
}

names(predictions) <- gsub(pattern = ".rds", replacement = "", x = names(predictions))
names(predictions) <- gsub(pattern = "mussel_", replacement = "", x = names(predictions))

#join the predictions to the list of comids
predictions <- cbind(covariates, predictions)

huc10 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU10/WBDHU10_NE.shp")
huc10 <- huc10 %>% 
  select(TNMID, AreaSqKm , Name)%>% 
  rename(huc10_areasqkm = AreaSqKm, huc10_name= Name, huc10_tnmid = TNMID)
huc10 <- huc10[!duplicated(huc10$huc10_tnmid),]#Remove the duplicated TNMID fields

predictions <- left_join(huc10, predictions, by = c("huc10_name", "huc10_tnmid"))


#sample plot
shp <- left_join(mussel_event_huc_join, mussel_occurrence, by = "UID") %>% 
  filter(!is.na(live_occurrence)) %>% 
  filter(live_occurrence == 1) %>% 
  filter(huc10_name != "Long Island Sound")
predictions <- predictions %>% 
  filter(huc10_name != "Long Island Sound")


  
ggplot(data = predictions)+
  geom_sf(aes(fill = `eastern lampmussel`))+
  scale_fill_gradient2(low = "grey60",
                        mid = "yellow",
                        high = "green3",
                        midpoint = 0.5,
                       limits = c(0,1))+
  geom_sf(data =shp[shp$common_name == "eastern lampmussel",], 
          aes(color = live_occurrence), color = "black", size = .5)

ggsave(paste("tmpfigures/baseline_mussel_predictions/easternlampmussel.png"), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300) 


```


apply the predictions to the climate change projections for flow and temperature

```{r}

#make the un-scaled baseline covariate file again
covariates <- left_join(NHDplusV2_NewEngCrop_mussel_covariates_HUC10, fish_host_prob_occ, by = c("huc10_name", "huc10_tnmid"))

covariates <- left_join(covariates, fish_host_propabun, by = c("huc10_name", "huc10_tnmid"))


#load the 2080 file that has the projected stream temperature and streamflow covariate values
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates_2080.RData")


#average by huc10 all the projected flow and temperature variables
test <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(huc10_name, huc10_tnmid, annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST) %>% 
  group_by(huc10_name, huc10_tnmid) %>% 
  summarise(annual_mean_summer_temp = mean(annual_mean_summer_temp, na.rm = T),
            BFI_HIST = mean(BFI_HIST, na.rm = T),
            LO7Q1DT_HIST = mean(LO7Q1DT_HIST, na.rm = T),
            CFM_HIST = mean(CFM_HIST, na.rm = T),
            W95_HIST = mean(W95_HIST, na.rm = T),
            logMJJA_HIST = mean(logMJJA_HIST, na.rm = T))

#need to order the new dataframe correctly and scale using the means and sd from the baseline data

#remove the current flow and temperature values from the baseline file and add in the 2080 projections
order <- names(covariates)
covariates2 <- covariates %>% 
  select(-c(annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST)) %>% #remove the baseline values for these variables
  left_join(test, by = c("huc10_name", "huc10_tnmid")) %>%  #add in the projected 2080 values
  select(all_of(order))


#scale the variables using the means and sd used to scale the variables in the baseline condition
test <- covariates2[,c(3:45)]
test <- sweep(test, 2, means, FUN = '-') 
test <- sweep(test, 2, sd, FUN = '/')  

#join the scaled variables back to the full dataset
covariates_cc <- cbind((covariates2)[,c(1:2)], test)



#apply models to the climate change covariates
file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "mussel_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = predict(mdl, newdata = covariates_cc, type = "response"))
names(predictions)[1] <- file.list[i]




for (i in 2:12) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = covariates_cc, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  
  
}

names(predictions) <- gsub(pattern = ".rds", replacement = "", x = names(predictions))
names(predictions) <- gsub(pattern = "mussel_", replacement = "", x = names(predictions))

#join the predictions to the list of comids
prediction_cc <- cbind(covariates_cc, predictions)

huc10 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU10/WBDHU10_NE.shp")
huc10 <- huc10 %>% 
  select(TNMID, AreaSqKm , Name)%>% 
  rename(huc10_areasqkm = AreaSqKm, huc10_name= Name, huc10_tnmid = TNMID)
huc10 <- huc10[!duplicated(huc10$huc10_tnmid),]#Remove the duplicated TNMID fields

prediction_cc <- left_join(huc10, prediction_cc, by = c("huc10_name", "huc10_tnmid"))




#sample plot
shp <- left_join(mussel_event_huc_join, mussel_occurrence, by = "UID") %>% 
  filter(!is.na(live_occurrence)) %>% 
  filter(live_occurrence == 1) %>% 
  filter(huc10_name != "Long Island Sound")
prediction_cc <- prediction_cc %>% 
  filter(huc10_name != "Long Island Sound")


  
ggplot(data = prediction_cc)+
  geom_sf(aes(fill = `yellow lampmussel`))+
  scale_fill_gradient2(low = "grey60",
                        mid = "yellow",
                        high = "green3",
                        midpoint = 0.5,
                       limits = c(0,1))+
  geom_sf(data =shp[shp$common_name == "yellow lampmussel",], 
          aes(color = live_occurrence), color = "black", size = .5)

ggsave(paste("tmpfigures/baseline_mussel_predictions/yellowlampmussel_cc.png"), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300) 

```

