---
title: "BiodiversityModelPrediction_Mussel_HUC12"
author: "Jenny Rogers"
date: "2023-04-14"
output: html_document
---

This is the BiodiversityModelPrediction_Mussel_HUC12 script, but updated to use the fish model results fro the BiodiversityModel_ClimateChangePaper script - ie its been adjusted to include riparian imperiousness(Ws), riparian natural(cat), and watershed forest % in the model, as opposed to watershed imperiousness.   This model will be used for the climate change and management paper (we are still using the original model in the paper 1)


In this script we model the probability of freshwater mussel occurrence at the HUC12 scale - I copied the script from "BiodiversityModel_Mussel.Rmd and just made changes so that everything was at the HUC12 scale instead of the HUC10 scale. The other difference is that the host fish variables have changed, so we model the different host fish predictions and average to HUC12

```{r}
library(tidyverse)
library(sf)
library(glmmTMB)


```


```{r}
#mussel data
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/mussel_occurrence.RData")
mussel_shp <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/sppdata/all_mussel_event.shp")

#mussel event and hydrography join
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/mussel_event_huc_join.RData")

#watershed shape files
huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")

#load New England state data
states <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/newenglandshape/NEWENGLAND_POLY.shp")


```

Baseline predictions


need to calculate the baseline host fish distributions at the HUC12 scale (we also did this in the biodiveristyModel_mussel_HUC12 file)

Host fish distributions at HUC12 scale.  This predicts the host fish distributions in the baseline condition. 
When we want to do mussel predictions in future time periods, we will need to redo this calcuation changing out the covariate data file for one with climate change and/or the management scenarios.  The new data sets need to be scaled according to the same mean and sd as the baseline.
```{r}

#load in covariates for the fish model and scale
load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates.RData")

#need to scal the variavles
test <- NHDplusV2_NewEngCrop %>% 
  data.frame() %>% 
  select(-geometry)

test <- test[,c(7:8, 10:42)]

#scale data
test <- data.frame(scale(test))

NHDplusV2_NewEngCrop <- cbind((NHDplusV2_NewEngCrop)[,c(1:6, 9)], test)







#Host fish variable 1 for the mussel model  - to be included in the model for all species of mussels:
#The predicted probability of occurrence of all fish (summed by stream reach – representing the total probability of occurrence of all species within a given reach) and averaged over a HUC12 – ultimately representing the average summed probability of occurrence of any fish species within a HUC12
#using the zprob from the prop abund model

file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop, type = "zprob")))) #do 1 - predict because we want prob of occurrence, not the prob of a zero
names(predictions)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop, type = "zprob"))))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  
  
}

colnames(predictions) <- gsub("_Mgmt_Paper_propabun_", "pres_", colnames(predictions)) #the probabun is the title of the model, but we used the zero predict option so we need to change the column names


#join the predictions to the list of comids
NHDplusV2_NewEngCrop_baseline <- cbind(NHDplusV2_NewEngCrop, predictions)

#first average by HUC12 and by species and save baseline fish predictions
fish_baseline_proboccurrence_huc12 <- NHDplusV2_NewEngCrop_baseline %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, 43:95) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(3:55, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, common_name) %>% 
  summarise(mean_prob_occ = mean(prob_occ)) %>% 
  ungroup() %>% 
  mutate(scenario = "baseline", taxa = "fish")
save(fish_baseline_proboccurrence_huc12, file = 'ProjectResults/fish_baseline_proboccurrence.RData')

#average by HUC10 becuase that is the scale we are modeling mussels
fish_host_prob_occ <- NHDplusV2_NewEngCrop_baseline %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, COMID, 43:95) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(4:56, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, COMID) %>% 
  summarise(total_prob_occ = sum(prob_occ)) %>% 
  ungroup() %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(mean_prob_occ = mean(total_prob_occ))
  






#fish host covariate with proportional abundance for the perferred host fish. (Variable 3 and 4)


file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop, type = "response"))
names(predictions)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  
  
}

#select only the columns for preferred host fish
predictions <- predictions %>% 
  select("_Mgmt_Paper_propabun_brown trout.rds", "_Mgmt_Paper_propabun_rainbow trout.rds", "_Mgmt_Paper_propabun_atlantic salmon.rds", "_Mgmt_Paper_propabun_brook trout.rds",
         "_Mgmt_Paper_propabun_yellow perch.rds", "_Mgmt_Paper_propabun_white perch.rds", "_Mgmt_Paper_propabun_largemouth bass.rds", "_Mgmt_Paper_propabun_bluegill.rds", "_Mgmt_Paper_propabun_pumpkinseed.rds", 
         "_Mgmt_Paper_propabun_redbreast sunfish.rds", )
names(predictions) <- gsub("_Mgmt_Paper_propabun_", "", names(predictions))
names(predictions) <- gsub(".rds", "", names(predictions))
names(predictions) <- gsub("\\.", " ", names(predictions))


#join the predictions to the list of comids
NHDplusV2_NewEngCrop_baseline <- cbind(NHDplusV2_NewEngCrop, predictions)

#average by HUC10 becuase that is the scale we are modeling mussels
fish_host_propabun <- NHDplusV2_NewEngCrop_baseline %>%
  data.frame() %>% 
  select(COMID, huc12_name, huc12_tnmid, 43:52) %>% 
  filter(!is.na(brook.trout),
         !is.na(huc12_name)) %>%
  mutate(sumYLamp = white.perch + yellow.perch,
         sumEPear = brown.trout + rainbow.trout + atlantic.salmon + brook.trout,
         sumEpond = yellow.perch + largemouth.bass + bluegill + pumpkinseed + redbreast.sunfish) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(TMuck_perferredhost = mean(white.perch),
            YLamp_perferredhost = mean(sumYLamp),
            EPear_perferredhost = mean(sumEPear),
            Epond_perferredhost = mean(sumEpond))




#Variable at the HUC12 scale showing if the non-modeled host fish has occurred in that HUC12 (Variable 2)
#fish data
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fish_event_huc_join.RData")
load('C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_presence.RData')

df <- left_join(fish_presence, fish_event_huc_join, by = "UID") %>% 
  filter(stock == "natural") %>% 
  select(huc12_name, huc12_tnmid, common_name) %>% 
  unique() %>% 
  filter(common_name %in% c("striped bass", 
                            "rainbow smelt", "american shad", "alewife", "blueback herring",
                            "ninespine stickleback", "three-spined stickleback", "fourspine stickleback"))



huc12 <- huc12 %>% 
  data.frame() %>% 
  select(name, tnmid) %>%
  rename(huc12_name = name, huc12_tnmid = tnmid) %>% 
  unique()
fish <- unique(df$common_name)

#make a dataframe that has each UID repeated for each fish spp
absence <- data.frame('huc12_name' = rep(huc12$huc12_name, each = 8,),
                      'huc12_tnmid' = rep(huc12$huc12_tnmid, each = 8))
absence$common_name <- rep(fish, times = nrow(huc12))
#keep only rows in the absence list that are no in the fish_presence list
keep <- anti_join(absence, df, by = c("huc12_name", "huc12_tnmid", "common_name"))

#add presence absence information
keep$occurrence <- 0
df$occurrence <- 1



#bind the presence and absence columns together
fish_host_occ <- rbind(df, keep)
fish_host_occ <- fish_host_occ %>% 
  arrange(huc12_name, huc12_tnmid) %>% 
  pivot_wider(names_from = common_name, values_from = occurrence)

rm(absence, fish_presence, keep)
rm(fish, huc12)

fish_host_occ$YLamp_notmodeledhost <- fish_host_occ$`striped bass`
fish_host_occ$Epond_notmodeledhost <- ifelse(fish_host_occ$`three-spined stickleback` == 1 |
                                                 fish_host_occ$`fourspine stickleback` == 1 |
                                                 fish_host_occ$`ninespine stickleback` == 1,
                                               1, 0)
fish_host_occ$Afloa_notmodeledhost <- ifelse(fish_host_occ$alewife == 1 |
                                                 fish_host_occ$`blueback herring` == 1 |
                                                 fish_host_occ$`american shad` == 1 |
                                                 fish_host_occ$`rainbow smelt` == 1,
                                               1, 0)


fish_host_occ <- fish_host_occ %>% 
  select(huc12_name, huc12_tnmid, YLamp_notmodeledhost, Epond_notmodeledhost, Afloa_notmodeledhost)





#plot fish host prob of occurrence
huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  rename("huc12_name" = "name",
         "huc12_tnmid" = "tnmid")
plot <- left_join(huc12, fish_host_propabun, by = c("huc12_name", "huc12_tnmid"))%>% 
  left_join(fish_host_occ, by = c("huc12_name", "huc12_tnmid"))%>% 
  left_join(fish_host_prob_occ, by = c("huc12_name", "huc12_tnmid"))%>% 
  filter(huc12_name != "Long Island Sound")

  
ggplot(data = plot)+
  geom_sf(aes(fill = YLamp_notmodeledhost), color = NA)+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank())+
  labs(fill = "Epond_notmodeledhost")

ggsave(paste("tmpfigures/Afloa_notmodeledhost.png"), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300)  


```




Load in the HUC12 mussel covariates,  join the host fish distributions, and scale
```{r}

load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_mussel_covariates_HUC12.RData")

covariates <- left_join(NHDplusV2_NewEngCrop_mussel_covariates_HUC12, fish_host_prob_occ, by = c("huc12_name", "huc12_tnmid"))

covariates <- left_join(covariates, fish_host_propabun, by = c("huc12_name", "huc12_tnmid"))

covariates <- left_join(covariates, fish_host_occ, by = c("huc12_name", "huc12_tnmid"))


#scale mussel covariates
test <- covariates[,c(3:42)]

means <- sapply(test, mean, na.rm = T)
save(means, file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_mean.RData")
sd <- sapply(test, sd, na.rm = T)
save(sd, file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_sd.RData")


#scale data
test <- data.frame(scale(test))

covariates <- cbind((covariates)[,c(1:2)], test)





```

Load in the mussel models and apply them to the baseline data

```{r}

file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "mussel_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = predict(mdl, newdata = covariates, type = "response"))
names(predictions)[1] <- file.list[i]




for (i in 2:12) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = covariates, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  
  
}

names(predictions) <- gsub(pattern = ".rds", replacement = "", x = names(predictions))
names(predictions) <- gsub(pattern = "mussel_", replacement = "", x = names(predictions))

#join the predictions to the list of comids
predictions <- cbind(covariates, predictions)

huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  select(tnmid, areasqkm , name)%>% 
  rename(huc12_areasqkm = areasqkm, huc12_name= name, huc12_tnmid = tnmid)
huc12 <- huc12[!duplicated(huc12$huc12_tnmid),]#Remove the duplicated TNMID fields

predictions_baseline <- left_join(huc12, predictions, by = c("huc12_name", "huc12_tnmid"))


#sample plot
shp <- left_join(mussel_event_huc_join, mussel_occurrence, by = "UID") %>% 
  filter(!is.na(live_occurrence)) %>% 
  filter(live_occurrence == 1) 
 


plots <- predictions_baseline %>% 
  select(44:55) %>% 
  rename("alewife floater" = "alewifefloater", 
         "eastern pearlshell" = "easternpearlshell", 
         "eastern pondmussel" = "easternpondmussel",
         "tidewater mucket" = "Tidewatermucket",
         "yellow lampmussel" = "yellowlampmussel")
for (i in (1:length(names(plots)))) { 

ggplot(data = plots)+
  geom_sf(aes(fill = plots[[i]]), color = NA)+
  scale_fill_gradient2(low = "grey60",
                        mid = "yellow",
                        high = "green3",
                        midpoint = 0.5,
                       limits = c(0,1))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        strip.background = element_rect(fill = "grey50",
                                        colour = "black"),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 15))+
    ggtitle(label = names(plots)[i])+
  labs(fill = "Probability")+
  geom_sf(data =shp[shp$common_name == names(plots)[i],], 
          aes(color = live_occurrence), color = "black", size = .5)

ggsave(paste("tmpfigures/baseline_mussel_predictions/", names(plots)[i], ".png", sep = ""), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300) 

print(i)
}




```
 
 CLIMATE CHANGE 
 mussel probability of occurrence with the climate change scenario for 2080
 first, re-calculate the host fish covariates with the climate change data
```{r}


#load the 2080 file that has the projected stream temperature and streamflow covariate values
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates_2080.RData")


#need to get the variables named the same as the baseline dataframe and need to order the variables in the same way so that we can scale them
test <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  select(7:8, 10:42) %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(1:2, 30:34, 3:27, 35, 28:29)

#load means and sd of the baseline covariates so we can scale the future data appropriates
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_mean.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_sd.RData")



#scale the variables using the means and sd used to scale the variables in the baseline condition
test <- sweep(test, 2, means, FUN = '-') 
test <- sweep(test, 2, sd, FUN = '/')  

#join the scaled variables back to 
NHDplusV2_NewEngCrop_cc <- cbind((NHDplusV2_NewEngCrop_2080)[,c(1:6, 9)], test)



#Host fish variable 1 for the mussel model  - to be included in the model for all species of mussels:
#The predicted probability of occurrence of all fish (summed by stream reach – representing the total probability of occurrence of all species within a given reach) and averaged over a HUC12 – ultimately representing the average summed probability of occurrence of any fish species within a HUC12
#using the zprob from the prop abund model

file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions_cc <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop_cc, type = "zprob")))) #do 1 - predict because we want prob of occurrence, not the prob of a zero
names(predictions_cc)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop_cc, type = "zprob"))))
  names(tmp)[1] <- file.list[i]
  
  predictions_cc <- cbind(predictions_cc, tmp)
  
  
}

colnames(predictions_cc) <- gsub("_Mgmt_Paper_propabun_", "pres_", colnames(predictions_cc)) #the probabun is the title of the model, but we used the zero predict option so we need to change the column names


#join the predictions_cc to the list of comids
tmp <- cbind(NHDplusV2_NewEngCrop_cc, predictions_cc)


#first average by HUC12 and by species and save cc fish predictions
fish_cc_proboccurrence_huc12 <- tmp %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, 43:95) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(3:55, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, common_name) %>% 
  summarise(mean_prob_occ = mean(prob_occ)) %>% 
  ungroup() %>% 
  mutate(scenario = "cc", taxa = "fish")
save(fish_cc_proboccurrence_huc12, file = 'ProjectResults/fish_cc_proboccurrence.RData')


#average by HUC12 becuase that is the scale we are modeling mussels
fish_host_prob_occ_cc <- tmp %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, COMID, 40:92) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(4:56, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, COMID) %>% 
  summarise(total_prob_occ = sum(prob_occ)) %>% 
  ungroup() %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(mean_prob_occ = mean(total_prob_occ))
  






#fish host covariate with proportional abundance for the perferred host fish. (Variable 3 and 4)


file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions_cc <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop_cc, type = "response"))
names(predictions_cc)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop_cc, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions_cc <- cbind(predictions_cc, tmp)
  
  
}

#select only the columns for preferred host fish
predictions_cc <- predictions_cc %>% 
  select("_Mgmt_Paper_propabun_brown trout.rds", "_Mgmt_Paper_propabun_rainbow trout.rds", "_Mgmt_Paper_propabun_atlantic salmon.rds", "_Mgmt_Paper_propabun_brook trout.rds",
         "_Mgmt_Paper_propabun_yellow perch.rds", "_Mgmt_Paper_propabun_white perch.rds", "_Mgmt_Paper_propabun_largemouth bass.rds", "_Mgmt_Paper_propabun_bluegill.rds", "_Mgmt_Paper_propabun_pumpkinseed.rds", 
         "_Mgmt_Paper_propabun_redbreast sunfish.rds", )
names(predictions_cc) <- gsub("_Mgmt_Paper_propabun_", "", names(predictions_cc))
names(predictions_cc) <- gsub(".rds", "", names(predictions_cc))
#names(predictions_cc) <- gsub("\\.", " ", names(predictions_cc))


#join the predictions to the list of comids
tmp <- cbind(NHDplusV2_NewEngCrop_cc, predictions_cc)

#average by HUC10 becuase that is the scale we are modeling mussels
fish_host_propabun_cc <- tmp %>%
  data.frame() %>% 
  select(COMID, huc12_name, huc12_tnmid, 43:52) %>% 
  filter(!is.na(brook.trout),
         !is.na(huc12_name)) %>%
  mutate(sumYLamp = white.perch + yellow.perch,
         sumEPear = brown.trout + rainbow.trout + atlantic.salmon + brook.trout,
         sumEpond = yellow.perch + largemouth.bass + bluegill + pumpkinseed + redbreast.sunfish) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(TMuck_perferredhost = mean(white.perch),
            YLamp_perferredhost = mean(sumYLamp),
            EPear_perferredhost = mean(sumEPear),
            Epond_perferredhost = mean(sumEpond))




#Variable at the HUC12 scale showing if the non-modeled host fish has occurred in that HUC12 (Variable 2)
#fish data
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fish_event_huc_join.RData")
load('C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_presence.RData')

fish_event_huc_join <- fish_event_huc_join %>% 
  data.frame() %>% 
  select(-geometry)

df <- left_join(fish_presence, fish_event_huc_join, by = "UID") %>% 
  filter(stock == "natural") %>% 
  select(huc12_name, huc12_tnmid, common_name) %>% 
  unique() %>% 
  filter(common_name %in% c("striped bass", 
                            "rainbow smelt", "american shad", "alewife", "blueback herring",
                            "ninespine stickleback", "three-spined stickleback", "fourspine stickleback"))


huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  data.frame() %>% 
  select(name, tnmid) %>%
  rename(huc12_name = name, huc12_tnmid = tnmid) %>% 
  unique()
fish <- unique(df$common_name)

#make a dataframe that has each UID repeated for each fish spp
absence <- data.frame('huc12_name' = rep(huc12$huc12_name, each = 8,),
                      'huc12_tnmid' = rep(huc12$huc12_tnmid, each = 8))
absence$common_name <- rep(fish, times = nrow(huc12))
#keep only rows in the absence list that are no in the fish_presence list
keep <- anti_join(absence, df, by = c("huc12_name", "huc12_tnmid", "common_name"))

#add presence absence information
keep$occurrence <- 0
df$occurrence <- 1



#bind the presence and absence columns together
fish_host_occ <- rbind(df, keep)
fish_host_occ <- fish_host_occ %>% 
  arrange(huc12_name, huc12_tnmid) %>% 
  pivot_wider(names_from = common_name, values_from = occurrence)

rm(absence, fish_presence, keep)
rm(fish, huc12)

fish_host_occ$YLamp_notmodeledhost <- fish_host_occ$`striped bass`
fish_host_occ$Epond_notmodeledhost <- ifelse(fish_host_occ$`three-spined stickleback` == 1 |
                                                 fish_host_occ$`fourspine stickleback` == 1 |
                                                 fish_host_occ$`ninespine stickleback` == 1,
                                               1, 0)
fish_host_occ$Afloa_notmodeledhost <- ifelse(fish_host_occ$alewife == 1 |
                                                 fish_host_occ$`blueback herring` == 1 |
                                                 fish_host_occ$`american shad` == 1 |
                                                 fish_host_occ$`rainbow smelt` == 1,
                                               1, 0)


fish_host_occ <- fish_host_occ %>% 
  select(huc12_name, huc12_tnmid, YLamp_notmodeledhost, Epond_notmodeledhost, Afloa_notmodeledhost)




#the correct host fish files for the climate change scenario include: fish_host_occ, fish_host_prob_occ_cc, fish_host_propabun_cc

```







apply the predictions to the climate change projections for flow and temperature


```{r}

load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_mussel_covariates_HUC12.RData")

covariates_cc <- left_join(NHDplusV2_NewEngCrop_mussel_covariates_HUC12, fish_host_prob_occ_cc, by = c("huc12_name", "huc12_tnmid"))

covariates_cc <- left_join(covariates_cc, fish_host_propabun_cc, by = c("huc12_name", "huc12_tnmid"))

covariates_cc <- left_join(covariates_cc, fish_host_occ, by = c("huc12_name", "huc12_tnmid"))


#load the 2080 file that has the projected stream temperature and streamflow covariate values
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates_2080.RData")


#average by huc12 all the projected flow and temperature variables
test <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(huc12_name, huc12_tnmid, annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(annual_mean_summer_temp = mean(annual_mean_summer_temp, na.rm = T),
            BFI_HIST = mean(BFI_HIST, na.rm = T),
            LO7Q1DT_HIST = mean(LO7Q1DT_HIST, na.rm = T),
            CFM_HIST = mean(CFM_HIST, na.rm = T),
            W95_HIST = mean(W95_HIST, na.rm = T),
            logMJJA_HIST = mean(logMJJA_HIST, na.rm = T))



#need to order the new dataframe correctly and scale using the means and sd from the baseline data

#remove the current flow and temperature values from the baseline file and add in the 2080 projections
order <- names(covariates)
covariates_cc2 <- covariates_cc %>% 
  select(-c(annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST)) %>% #remove the baseline values for these variables
  left_join(test, by = c("huc12_name", "huc12_tnmid")) %>%  #add in the projected 2080 values
  select(all_of(order))

#load the values used to scale the mussel covariates
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_mean.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_sd.RData")



#scale the variables using the means and sd used to scale the variables in the baseline condition
test <- covariates_cc2[,c(3:42)]
test <- sweep(test, 2, means, FUN = '-') 
test <- sweep(test, 2, sd, FUN = '/')  

#join the scaled variables back to the full dataset
covariates_cc <- cbind((covariates_cc2)[,c(1:2)], test)



#apply models to the climate change covariates
file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "mussel_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = predict(mdl, newdata = covariates_cc, type = "response"))
names(predictions)[1] <- file.list[i]




for (i in 2:12) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = covariates_cc, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  
  
}

names(predictions) <- gsub(pattern = ".rds", replacement = "", x = names(predictions))
names(predictions) <- gsub(pattern = "mussel_", replacement = "", x = names(predictions))

#join the predictions to the list of comids
prediction_cc <- cbind(covariates_cc, predictions)

huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  select(tnmid, areasqkm , name)%>% 
  rename(huc12_areasqkm = areasqkm, huc12_name= name, huc12_tnmid = tnmid)
huc12 <- huc12[!duplicated(huc12$huc12_tnmid),]#Remove the duplicated TNMID fields

prediction_cc <- left_join(huc12, prediction_cc, by = c("huc12_name", "huc12_tnmid"))




#sample plot
shp <- left_join(mussel_event_huc_join, mussel_occurrence, by = "UID") %>% 
  filter(!is.na(live_occurrence)) %>% 
  filter(live_occurrence == 1) 



plots <- prediction_cc %>% 
  select(44:55) %>% 
  rename("alewife floater" = "alewifefloater", 
         "eastern pearlshell" = "easternpearlshell", 
         "eastern pondmussel" = "easternpondmussel",
         "tidewater mucket" = "Tidewatermucket",
         "yellow lampmussel" = "yellowlampmussel")

for (i in (1:length(names(plots)))) { 

ggplot(data = plots)+
  geom_sf(aes(fill = plots[[i]]), color = NA)+
  scale_fill_gradient2(low = "grey60",
                        mid = "yellow",
                        high = "green3",
                        midpoint = 0.5,
                       limits = c(0,1))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        strip.background = element_rect(fill = "grey50",
                                        colour = "black"),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 15))+
    ggtitle(label = names(plots)[i])+
  labs(fill = "Probability")+
  geom_sf(data =shp[shp$common_name == names(plots)[i],], 
          aes(color = live_occurrence), color = "black", size = .5)

ggsave(paste("tmpfigures/baseline_mussel_predictions/", names(plots)[i], "_cc.png", sep = ""), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300) 

print(i)
}

```





species richness, change in species richness, change for each species
```{r}
#change from baseline to climate change for each species

prediction_cc <- data.frame(prediction_cc)
predictions_baseline <- data.frame(predictions_baseline)

#save data files with mussel predictions for baseline and climate change
save(prediction_cc, file = 'ProjectResults/mussel_CC2080_proboccurrence.RData')
save(predictions_baseline, file = 'ProjectResults/mussel_baseline_proboccurrence.RData')

cc <- as.matrix(prediction_cc[, 44:55])
bl <- as.matrix(predictions_baseline[, 44:55])

change <- (cc - bl)
change <- data.frame(change)
change <- cbind(predictions_baseline[1:3], change)

change <- left_join(huc12, change, by = c("huc12_name", "huc12_tnmid", "huc12_areasqkm"))

names(change) <- gsub("\\.", " ", names(change))


#sample plot
shp <- left_join(mussel_event_huc_join, mussel_occurrence, by = "UID") %>% 
  filter(!is.na(live_occurrence)) %>% 
  filter(live_occurrence == 1) %>% 
  filter(huc12_name != "Long Island Sound Deep")
change <- change %>% 
  filter(huc12_name != "Long Island Sound Deep")


  
ggplot(data = change)+
  geom_sf(aes(fill = `eastern floater`), linewidth = .02)+
  scale_fill_gradient2(low = "red2",
                        mid = "white",
                        high = "green4",
                        midpoint = 0,
                       limits = c(-1,1))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        strip.background = element_rect(fill = "grey50",
                                        colour = "black"),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))+
    ggtitle(label = "eastern floater")+
  labs(fill = "Δ P")

ggsave(paste("tmpfigures/baseline_mussel_predictions/tidewatermucket_change.png"), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300) 



#species richness in the baseline
richness_baseline <- predictions_baseline %>% 
  select(1,3,44:55) %>% 
  pivot_longer(3:14, names_to = "common_name", values_to = "prob") %>% 
  mutate(occurrence = ifelse(common_name %in% c("alewifefloater", "brook.floater", "dwarf.wedgemussel", "easternpondmussel", "Tidewatermucket", "yellowlampmussel") &
                                                  prob > 0.1 , 1,
                                                ifelse(common_name %in% c("creeper", "eastern.elliptio", "eastern.floater", "eastern.lampmussel", "easternpearlshell", "triangle.floater") & prob > 0.5, 1, 0)))  %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(richness = sum(occurrence))


baseline <- left_join(huc12, richness_baseline, by = c("huc12_name", "huc12_tnmid"))%>% 
  filter(huc12_name != "Long Island Sound Deep")

  
ggplot(data = baseline)+
  geom_sf(aes(fill = richness))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))+
    ggtitle(label = "Baseline Species Richness")+
  labs(fill = "Richness")

ggsave(paste("tmpfigures/baseline_mussel_predictions/mussel_richness_baseline.png"), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300) 





#species richness in climate change
richness_cc <- prediction_cc %>% 
  select(1,3,44:55) %>% 
  pivot_longer(3:14, names_to = "common_name", values_to = "prob") %>% 
  mutate(occurrence = ifelse(common_name %in% c("alewifefloater", "brook.floater", "dwarf.wedgemussel", "easternpondmussel", "Tidewatermucket", "yellowlampmussel") &
                                                  prob > 0.1 , 1,
                                                ifelse(common_name %in% c("creeper", "eastern.elliptio", "eastern.floater", "eastern.lampmussel", "easternpearlshell", "triangle.floater") & prob > 0.5, 1, 0)))  %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(richness = sum(occurrence))






#change in species richness: climate change - baseline

change_richness <- data.frame(
  "huc12_tnmid" = richness_baseline[,1],
  "huc12_name" = richness_baseline[,2],
  "richness_change" = richness_cc$richness - richness_baseline$richness
)


change_richness <- left_join(huc12, change_richness, by = c("huc12_name", "huc12_tnmid"))%>% 
  filter(huc12_name != "Long Island Sound Deep")

  
ggplot(data = change_richness)+
  geom_sf(aes(fill = richness_change))+
  scale_fill_gradient2(low = "red2",
                        mid = "white",
                        high = "green4",
                        midpoint = 0,
                       limits = c(min(change_richness$richness_change, na.rm = T), max(change_richness$richness_change, na.rm = T)))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))+
    ggtitle(label = "Change Species Richness")+
  labs(fill = "	Δ Richness")

ggsave(paste("tmpfigures/baseline_mussel_predictions/mussel_richness_change.png"), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300) 





#align change in richess values with change in the climate change varaibles
covariates_bl <- NHDplusV2_NewEngCrop_mussel_covariates_HUC12 %>% 
  select(huc12_name, huc12_tnmid, annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST)

#Climate change streamflow and temperature variables
covariates_cc <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(huc12_name, huc12_tnmid, annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(annual_mean_summer_temp = mean(annual_mean_summer_temp, na.rm = T),
            BFI_HIST = mean(BFI_HIST, na.rm = T),
            LO7Q1DT_HIST = mean(LO7Q1DT_HIST, na.rm = T),
            CFM_HIST = mean(CFM_HIST, na.rm = T),
            W95_HIST = mean(W95_HIST, na.rm = T),
            logMJJA_HIST = mean(logMJJA_HIST, na.rm = T))


sum(names(covariates_bl) == names(covariates_cc)) #45
sum(covariates_bl$huc12_name == covariates_cc$huc12_name, na.rm = T) #423

bl_cov <- covariates_bl %>% 
  ungroup() %>% 
  select(annual_mean_summer_temp, W95_HIST, logMJJA_HIST) %>% 
  mutate(mjja_hist = exp(logMJJA_HIST)) %>% 
  as.matrix()

cc_cov <- covariates_cc %>% 
  ungroup() %>% 
  select(annual_mean_summer_temp, W95_HIST, logMJJA_HIST) %>% 
  mutate(mjja_hist = exp(logMJJA_HIST)) %>% 
  as.matrix()

change_cov <- (cc_cov - bl_cov)
change_cov <- data.frame(change_cov)
change_cov <- cbind(covariates_bl[1:2], change_cov)


change_cov <- left_join(huc12, change_cov, by = c("huc12_name", "huc12_tnmid"))%>% 
  filter(huc12_name != "Long Island Sound Deep")

  
ggplot(data = change_cov)+
  geom_sf(aes(fill = annual_mean_summer_temp))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 20))+
    ggtitle(label = "Annual mean summer temp Change")+
  labs(fill = "Δ T(°C)")

ggsave(paste("tmpfigures/baseline_mussel_predictions/temp_change.png"), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300)  


change_richness <- change_richness %>% 
  data.frame() %>% 
  select(huc12_tnmid, huc12_name, richness_change)
df <- left_join(change_cov, change_richness, by = c("huc10_name", "huc10_tnmid"))

df <- df %>% 
  filter(!is.na(annual_mean_summer_temp))

r2 <- round(cor(df$richness_change, df$annual_mean_summer_temp),2)
cor.test(df$richness_change, df$W95_HIST)
label <- paste("R2 = ",  r2, sep = "")

ggplot(data = df, mapping = aes(x = annual_mean_summer_temp, y = richness_change))+
  geom_jitter(alpha = 0.1, size = 3)+
  geom_smooth(method = "lm", linewidth = 2, color = "red")+
  labs(y = "Change in Richness",
       x = "Change in Summer Stream Temperature (°C)")+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.title = element_text(size = 17),
          axis.text = element_text(size = 17))+
  annotate("text",x=3,y=-3,label=r2 ,parse=TRUE, color = "red", size = 5)



```



MANAGEMENT AND CLIMATE CHANGE: PctImp_WsRp100 = 0 

```{r}


#load the 2080 file that has the projected stream temperature and streamflow covariate values
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates_2080.RData")


#need to get the variables named the same as the baseline dataframe and need to order the variables in the same way so that we can scale them
test <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  select(7:8, 10:42) %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(1:2, 30:34, 3:27, 35, 28:29)


#management scenario - decrease PctImp_WsRp100 

test$PctImp_WsRp100 <- 0


#load means and sd of the baseline covariates so we can scale the future data appropriates
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_mean.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_sd.RData")



#scale the variables using the means and sd used to scale the variables in the baseline condition
test <- sweep(test, 2, means, FUN = '-') 
test <- sweep(test, 2, sd, FUN = '/')  

#join the scaled variables back to 
NHDplusV2_NewEngCrop_impervious <- cbind((NHDplusV2_NewEngCrop_2080)[,c(1:6, 9)], test)



#Host fish variable 1 for the mussel model  - to be included in the model for all species of mussels:
#The predicted probability of occurrence of all fish (summed by stream reach – representing the total probability of occurrence of all species within a given reach) and averaged over a HUC12 – ultimately representing the average summed probability of occurrence of any fish species within a HUC12
#using the zprob from the prop abund model

file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions_impervious <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop_impervious, type = "zprob")))) #do 1 - predict because we want prob of occurrence, not the prob of a zero
names(predictions_impervious)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop_impervious, type = "zprob"))))
  names(tmp)[1] <- file.list[i]
  
  predictions_impervious <- cbind(predictions_impervious, tmp)
  
  
}

colnames(predictions_impervious) <- gsub("_Mgmt_Paper_propabun_", "pres_", colnames(predictions_impervious)) #the probabun is the title of the model, but we used the zero predict option so we need to change the column names


#join the predictions_impervious to the list of comids
tmp <- cbind(NHDplusV2_NewEngCrop_impervious, predictions_impervious)


#first average by HUC12 and by species and save cc fish predictions
fish_impervious_proboccurrence_huc12 <- tmp %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, 43:95) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(3:55, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, common_name) %>% 
  summarise(mean_prob_occ = mean(prob_occ)) %>% 
  ungroup() %>% 
  mutate(scenario = "impervious", taxa = "fish")
save(fish_impervious_proboccurrence_huc12, file = 'ProjectResults/fish_impervious_proboccurrence.RData')


#average by HUC12 becuase that is the scale we are modeling mussels
fish_host_prob_occ_impervious <- tmp %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, COMID, 43:95) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(4:56, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, COMID) %>% 
  summarise(total_prob_occ = sum(prob_occ)) %>% 
  ungroup() %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(mean_prob_occ = mean(total_prob_occ))
  






#fish host covariate with proportional abundance for the perferred host fish. (Variable 3 and 4)


file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions_impervious <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop_impervious, type = "response"))
names(predictions_impervious)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop_impervious, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions_impervious <- cbind(predictions_impervious, tmp)
  
  
}

#select only the columns for preferred host fish
predictions_impervious <- predictions_impervious %>% 
  select("_Mgmt_Paper_propabun_brown trout.rds", "_Mgmt_Paper_propabun_rainbow trout.rds", "_Mgmt_Paper_propabun_atlantic salmon.rds", "_Mgmt_Paper_propabun_brook trout.rds",
         "_Mgmt_Paper_propabun_yellow perch.rds", "_Mgmt_Paper_propabun_white perch.rds", "_Mgmt_Paper_propabun_largemouth bass.rds", "_Mgmt_Paper_propabun_bluegill.rds", "_Mgmt_Paper_propabun_pumpkinseed.rds", 
         "_Mgmt_Paper_propabun_redbreast sunfish.rds", )
names(predictions_impervious) <- gsub("_Mgmt_Paper_propabun_", "", names(predictions_impervious))
names(predictions_impervious) <- gsub(".rds", "", names(predictions_impervious))
#names(predictions_impervious) <- gsub("\\.", " ", names(predictions_impervious))


#join the predictions to the list of comids
tmp <- cbind(NHDplusV2_NewEngCrop_impervious, predictions_impervious)

#average by HUC12 becuase that is the scale we are modeling mussels
fish_host_propabun_impervious <- tmp %>%
  data.frame() %>% 
  select(COMID, huc12_name, huc12_tnmid, 43:52) %>% 
  filter(!is.na(brook.trout),
         !is.na(huc12_name)) %>%
  mutate(sumYLamp = white.perch + yellow.perch,
         sumEPear = brown.trout + rainbow.trout + atlantic.salmon + brook.trout,
         sumEpond = yellow.perch + largemouth.bass + bluegill + pumpkinseed + redbreast.sunfish) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(TMuck_perferredhost = mean(white.perch),
            YLamp_perferredhost = mean(sumYLamp),
            EPear_perferredhost = mean(sumEPear),
            Epond_perferredhost = mean(sumEpond))




#Variable at the HUC12 scale showing if the non-modeled host fish has occurred in that HUC12 (Variable 2)
#fish data
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fish_event_huc_join.RData")
load('C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_presence.RData')

fish_event_huc_join <- fish_event_huc_join %>% 
  data.frame() %>% 
  select(-geometry)

df <- left_join(fish_presence, fish_event_huc_join, by = "UID") %>% 
  filter(stock == "natural") %>% 
  select(huc12_name, huc12_tnmid, common_name) %>% 
  unique() %>% 
  filter(common_name %in% c("striped bass", 
                            "rainbow smelt", "american shad", "alewife", "blueback herring",
                            "ninespine stickleback", "three-spined stickleback", "fourspine stickleback"))


huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  data.frame() %>% 
  select(name, tnmid) %>%
  rename(huc12_name = name, huc12_tnmid = tnmid) %>% 
  unique()
fish <- unique(df$common_name)

#make a dataframe that has each UID repeated for each fish spp
absence <- data.frame('huc12_name' = rep(huc12$huc12_name, each = 8,),
                      'huc12_tnmid' = rep(huc12$huc12_tnmid, each = 8))
absence$common_name <- rep(fish, times = nrow(huc12))
#keep only rows in the absence list that are no in the fish_presence list
keep <- anti_join(absence, df, by = c("huc12_name", "huc12_tnmid", "common_name"))

#add presence absence information
keep$occurrence <- 0
df$occurrence <- 1



#bind the presence and absence columns together
fish_host_occ <- rbind(df, keep)
fish_host_occ <- fish_host_occ %>% 
  arrange(huc12_name, huc12_tnmid) %>% 
  pivot_wider(names_from = common_name, values_from = occurrence)

rm(absence, fish_presence, keep)
rm(fish, huc12)

fish_host_occ$YLamp_notmodeledhost <- fish_host_occ$`striped bass`
fish_host_occ$Epond_notmodeledhost <- ifelse(fish_host_occ$`three-spined stickleback` == 1 |
                                                 fish_host_occ$`fourspine stickleback` == 1 |
                                                 fish_host_occ$`ninespine stickleback` == 1,
                                               1, 0)
fish_host_occ$Afloa_notmodeledhost <- ifelse(fish_host_occ$alewife == 1 |
                                                 fish_host_occ$`blueback herring` == 1 |
                                                 fish_host_occ$`american shad` == 1 |
                                                 fish_host_occ$`rainbow smelt` == 1,
                                               1, 0)


fish_host_occ <- fish_host_occ %>% 
  select(huc12_name, huc12_tnmid, YLamp_notmodeledhost, Epond_notmodeledhost, Afloa_notmodeledhost)




#the correct host fish files for the management scenario include: fish_host_occ, fish_host_prob_occ_cc, fish_host_propabun_cc

```


apply predictions to the climate change projections for flow AND temperature combined with the riparian imperiousness management


```{r}

load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_mussel_covariates_HUC12.RData")

covariates_impervious <- left_join(NHDplusV2_NewEngCrop_mussel_covariates_HUC12, fish_host_prob_occ_impervious, by = c("huc12_name", "huc12_tnmid"))

covariates_impervious <- left_join(covariates_impervious, fish_host_propabun_impervious, by = c("huc12_name", "huc12_tnmid"))

covariates_impervious <- left_join(covariates_impervious, fish_host_occ, by = c("huc12_name", "huc12_tnmid"))


#load the 2080 file that has the projected stream temperature and streamflow covariate values
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates_2080.RData")


#average by huc12 all the projected flow and temperature variables
test <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(huc12_name, huc12_tnmid, annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(annual_mean_summer_temp = mean(annual_mean_summer_temp, na.rm = T),
            BFI_HIST = mean(BFI_HIST, na.rm = T),
            LO7Q1DT_HIST = mean(LO7Q1DT_HIST, na.rm = T),
            CFM_HIST = mean(CFM_HIST, na.rm = T),
            W95_HIST = mean(W95_HIST, na.rm = T),
            logMJJA_HIST = mean(logMJJA_HIST, na.rm = T))


#need to order the new dataframe correctly and scale using the means and sd from the baseline data

#remove the current flow and temperature values from the baseline file and add in the 2080 projections
order <- names(covariates_impervious)
covariates_impervious2 <- covariates_impervious %>% 
  select(-c(annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST)) %>% #remove the baseline values for these variables
  left_join(test, by = c("huc12_name", "huc12_tnmid")) %>%  #add in the projected 2080 values
  select(all_of(order))

#management scenario - decrease PctImp_WsRp100 to 0

covariates_impervious2$PctImp_WsRp100 <- 0


#load the values used to scale the mussel covariates
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_mean.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_sd.RData")



#scale the variables using the means and sd used to scale the variables in the baseline condition
test <- covariates_impervious2[,c(3:42)]
test <- sweep(test, 2, means, FUN = '-') 
test <- sweep(test, 2, sd, FUN = '/')  

#join the scaled variables back to the full dataset
covariates_impervious <- cbind((covariates_impervious2)[,c(1:2)], test)



#apply models to the climate change and management covariates
file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "mussel_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = predict(mdl, newdata = covariates_impervious, type = "response"))
names(predictions)[1] <- file.list[i]




for (i in 2:12) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = covariates_impervious, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  
  
}

names(predictions) <- gsub(pattern = ".rds", replacement = "", x = names(predictions))
names(predictions) <- gsub(pattern = "mussel_", replacement = "", x = names(predictions))

#join the predictions to the list of comids
prediction_impervious <- cbind(covariates_impervious, predictions)

huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  select(tnmid, areasqkm , name)%>% 
  rename(huc12_areasqkm = areasqkm, huc12_name= name, huc12_tnmid = tnmid)
huc12 <- huc12[!duplicated(huc12$huc12_tnmid),]#Remove the duplicated TNMID fields

prediction_impervious <- left_join(huc12, prediction_impervious, by = c("huc12_name", "huc12_tnmid"))


#save data files with mussel predictions for climate change/landcover mgmt
save(prediction_impervious, file = 'ProjectResults/mussel_CC2080_landcover_impervious_proboccurrence.RData')


#sample plot
shp <- left_join(mussel_event_huc_join, mussel_occurrence, by = "UID") %>% 
  filter(!is.na(live_occurrence)) %>% 
  filter(live_occurrence == 1) %>% 
  filter(huc12_name != "Long Island Sound Deep")



  
plots <- prediction_impervious %>% 
  select(44:55) %>% 
  rename("alewife floater" = "alewifefloater", 
         "eastern pearlshell" = "easternpearlshell", 
         "eastern pondmussel" = "easternpondmussel",
         "tidewater mucket" = "Tidewatermucket",
         "yellow lampmussel" = "yellowlampmussel")

for (i in (1:length(names(plots)))) { 

ggplot(data = plots)+
  geom_sf(aes(fill = plots[[i]]), color = NA)+
  scale_fill_gradient2(low = "grey60",
                        mid = "yellow",
                        high = "green3",
                        midpoint = 0.5,
                       limits = c(0,1))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        strip.background = element_rect(fill = "grey50",
                                        colour = "black"),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 15))+
    ggtitle(label = names(plots)[i])+
  labs(fill = "Probability")+
  geom_sf(data =shp[shp$common_name == names(plots)[i],], 
          aes(color = live_occurrence), color = "black", size = .5)

ggsave(paste("tmpfigures/baseline_mussel_predictions/", names(plots)[i], "_impervious.png", sep = ""), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300) 

print(i)
}

```



MANAGEMENT AND CLIMATE CHANGE: pctNatural_ripbuf100 = 100 

```{r}

#load the 2080 file that has the projected stream temperature and streamflow covariate values
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates_2080.RData")


#need to get the variables named the same as the baseline dataframe and need to order the variables in the same way so that we can scale them
test <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  select(7:8, 10:42) %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(1:2, 30:34, 3:27, 35, 28:29)


#management scenario - increase pctNatural_ripbuf100 to 100

test$pctNatural_ripbuf100 <- 100

#load means and sd of the baseline covariates so we can scale the future data appropriates
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_mean.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_sd.RData")



#scale the variables using the means and sd used to scale the variables in the baseline condition
test <- sweep(test, 2, means, FUN = '-') 
test <- sweep(test, 2, sd, FUN = '/')  

#join the scaled variables back to 
NHDplusV2_NewEngCrop_natural <- cbind((NHDplusV2_NewEngCrop_2080)[,c(1:6, 9)], test)



#Host fish variable 1 for the mussel model  - to be included in the model for all species of mussels:
#The predicted probability of occurrence of all fish (summed by stream reach – representing the total probability of occurrence of all species within a given reach) and averaged over a HUC12 – ultimately representing the average summed probability of occurrence of any fish species within a HUC12
#using the zprob from the prop abund model

file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions_natural <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop_natural, type = "zprob")))) #do 1 - predict because we want prob of occurrence, not the prob of a zero
names(predictions_natural)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop_natural, type = "zprob"))))
  names(tmp)[1] <- file.list[i]
  
  predictions_natural <- cbind(predictions_natural, tmp)
  
  
}

colnames(predictions_natural) <- gsub("_Mgmt_Paper_propabun_", "pres_", colnames(predictions_natural)) #the probabun is the title of the model, but we used the zero predict option so we need to change the column names


#join the predictions_natural to the list of comids
tmp <- cbind(NHDplusV2_NewEngCrop_natural, predictions_natural)


#first average by HUC12 and by species and save cc fish predictions
fish_natural_proboccurrence_huc12 <- tmp %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, 43:95) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(3:55, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, common_name) %>% 
  summarise(mean_prob_occ = mean(prob_occ)) %>% 
  ungroup() %>% 
  mutate(scenario = "natural", taxa = "fish")
save(fish_natural_proboccurrence_huc12, file = 'ProjectResults/fish_natural_proboccurrence.RData')


#average by HUC12 becuase that is the scale we are modeling mussels
fish_host_prob_occ_natural <- tmp %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, COMID, 43:95) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(4:56, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, COMID) %>% 
  summarise(total_prob_occ = sum(prob_occ)) %>% 
  ungroup() %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(mean_prob_occ = mean(total_prob_occ))
  






#fish host covariate with proportional abundance for the perferred host fish. (Variable 3 and 4)


file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions_natural <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop_natural, type = "response"))
names(predictions_natural)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop_natural, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions_natural <- cbind(predictions_natural, tmp)
  
  
}

#select only the columns for preferred host fish
predictions_natural <- predictions_natural %>% 
  select("_Mgmt_Paper_propabun_brown trout.rds", "_Mgmt_Paper_propabun_rainbow trout.rds", "_Mgmt_Paper_propabun_atlantic salmon.rds", "_Mgmt_Paper_propabun_brook trout.rds",
         "_Mgmt_Paper_propabun_yellow perch.rds", "_Mgmt_Paper_propabun_white perch.rds", "_Mgmt_Paper_propabun_largemouth bass.rds", "_Mgmt_Paper_propabun_bluegill.rds", "_Mgmt_Paper_propabun_pumpkinseed.rds", 
         "_Mgmt_Paper_propabun_redbreast sunfish.rds", )
names(predictions_natural) <- gsub("_Mgmt_Paper_propabun_", "", names(predictions_natural))
names(predictions_natural) <- gsub(".rds", "", names(predictions_natural))
#names(predictions_natural) <- gsub("\\.", " ", names(predictions_natural))


#join the predictions to the list of comids
tmp <- cbind(NHDplusV2_NewEngCrop_natural, predictions_natural)

#average by HUC12 becuase that is the scale we are modeling mussels
fish_host_propabun_natural <- tmp %>%
  data.frame() %>% 
  select(COMID, huc12_name, huc12_tnmid, 43:52) %>% 
  filter(!is.na(brook.trout),
         !is.na(huc12_name)) %>%
  mutate(sumYLamp = white.perch + yellow.perch,
         sumEPear = brown.trout + rainbow.trout + atlantic.salmon + brook.trout,
         sumEpond = yellow.perch + largemouth.bass + bluegill + pumpkinseed + redbreast.sunfish) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(TMuck_perferredhost = mean(white.perch),
            YLamp_perferredhost = mean(sumYLamp),
            EPear_perferredhost = mean(sumEPear),
            Epond_perferredhost = mean(sumEpond))




#Variable at the HUC12 scale showing if the non-modeled host fish has occurred in that HUC12 (Variable 2)
#fish data
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fish_event_huc_join.RData")
load('C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_presence.RData')

fish_event_huc_join <- fish_event_huc_join %>% 
  data.frame() %>% 
  select(-geometry)

df <- left_join(fish_presence, fish_event_huc_join, by = "UID") %>% 
  filter(stock == "natural") %>% 
  select(huc12_name, huc12_tnmid, common_name) %>% 
  unique() %>% 
  filter(common_name %in% c("striped bass", 
                            "rainbow smelt", "american shad", "alewife", "blueback herring",
                            "ninespine stickleback", "three-spined stickleback", "fourspine stickleback"))


huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  data.frame() %>% 
  select(name, tnmid) %>%
  rename(huc12_name = name, huc12_tnmid = tnmid) %>% 
  unique()
fish <- unique(df$common_name)

#make a dataframe that has each UID repeated for each fish spp
absence <- data.frame('huc12_name' = rep(huc12$huc12_name, each = 8,),
                      'huc12_tnmid' = rep(huc12$huc12_tnmid, each = 8))
absence$common_name <- rep(fish, times = nrow(huc12))
#keep only rows in the absence list that are no in the fish_presence list
keep <- anti_join(absence, df, by = c("huc12_name", "huc12_tnmid", "common_name"))

#add presence absence information
keep$occurrence <- 0
df$occurrence <- 1



#bind the presence and absence columns together
fish_host_occ <- rbind(df, keep)
fish_host_occ <- fish_host_occ %>% 
  arrange(huc12_name, huc12_tnmid) %>% 
  pivot_wider(names_from = common_name, values_from = occurrence)

rm(absence, fish_presence, keep)
rm(fish, huc12)

fish_host_occ$YLamp_notmodeledhost <- fish_host_occ$`striped bass`
fish_host_occ$Epond_notmodeledhost <- ifelse(fish_host_occ$`three-spined stickleback` == 1 |
                                                 fish_host_occ$`fourspine stickleback` == 1 |
                                                 fish_host_occ$`ninespine stickleback` == 1,
                                               1, 0)
fish_host_occ$Afloa_notmodeledhost <- ifelse(fish_host_occ$alewife == 1 |
                                                 fish_host_occ$`blueback herring` == 1 |
                                                 fish_host_occ$`american shad` == 1 |
                                                 fish_host_occ$`rainbow smelt` == 1,
                                               1, 0)


fish_host_occ <- fish_host_occ %>% 
  select(huc12_name, huc12_tnmid, YLamp_notmodeledhost, Epond_notmodeledhost, Afloa_notmodeledhost)




#the correct host fish files for the management scenario include: fish_host_occ, fish_host_prob_occ_cc, fish_host_propabun_cc

```



apply the predictions to the climate change projections for flow AND temperature combined with the natural riparian buffer management 


```{r}

load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_mussel_covariates_HUC12.RData")

covariates_natural <- left_join(NHDplusV2_NewEngCrop_mussel_covariates_HUC12, fish_host_prob_occ_natural, by = c("huc12_name", "huc12_tnmid"))

covariates_natural <- left_join(covariates_natural, fish_host_propabun_natural, by = c("huc12_name", "huc12_tnmid"))

covariates_natural <- left_join(covariates_natural, fish_host_occ, by = c("huc12_name", "huc12_tnmid"))


#load the 2080 file that has the projected stream temperature and streamflow covariate values
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates_2080.RData")


#average by huc12 all the projected flow and temperature variables
test <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(huc12_name, huc12_tnmid, annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(annual_mean_summer_temp = mean(annual_mean_summer_temp, na.rm = T),
            BFI_HIST = mean(BFI_HIST, na.rm = T),
            LO7Q1DT_HIST = mean(LO7Q1DT_HIST, na.rm = T),
            CFM_HIST = mean(CFM_HIST, na.rm = T),
            W95_HIST = mean(W95_HIST, na.rm = T),
            logMJJA_HIST = mean(logMJJA_HIST, na.rm = T))


#need to order the new dataframe correctly and scale using the means and sd from the baseline data

#remove the current flow and temperature values from the baseline file and add in the 2080 projections
order <- names(covariates_natural)
covariates_natural2 <- covariates_natural %>% 
  select(-c(annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST)) %>% #remove the baseline values for these variables
  left_join(test, by = c("huc12_name", "huc12_tnmid")) %>%  #add in the projected 2080 values
  select(all_of(order))

#management scenario - increase pctNatural_ripbuf100 to 100

covariates_natural2$pctNatural_ripbuf100  <- 100


#load the values used to scale the mussel covariates
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_mean.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_sd.RData")



#scale the variables using the means and sd used to scale the variables in the baseline condition
test <- covariates_natural2[,c(3:42)]
test <- sweep(test, 2, means, FUN = '-') 
test <- sweep(test, 2, sd, FUN = '/')  

#join the scaled variables back to the full dataset
covariates_natural <- cbind((covariates_natural2)[,c(1:2)], test)



#apply models to the climate change and management covariates
file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "mussel_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = predict(mdl, newdata = covariates_natural, type = "response"))
names(predictions)[1] <- file.list[i]




for (i in 2:12) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = covariates_natural, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  
  
}

names(predictions) <- gsub(pattern = ".rds", replacement = "", x = names(predictions))
names(predictions) <- gsub(pattern = "mussel_", replacement = "", x = names(predictions))

#join the predictions to the list of comids
prediction_natural <- cbind(covariates_natural, predictions)

huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  select(tnmid, areasqkm , name)%>% 
  rename(huc12_areasqkm = areasqkm, huc12_name= name, huc12_tnmid = tnmid)
huc12 <- huc12[!duplicated(huc12$huc12_tnmid),]#Remove the duplicated TNMID fields

prediction_natural <- left_join(huc12, prediction_natural, by = c("huc12_name", "huc12_tnmid"))


#save data files with mussel predictions for climate change/landcover mgmt
save(prediction_natural, file = 'ProjectResults/mussel_CC2080_landcover_natural_proboccurrence.RData')


#sample plot
shp <- left_join(mussel_event_huc_join, mussel_occurrence, by = "UID") %>% 
  filter(!is.na(live_occurrence)) %>% 
  filter(live_occurrence == 1) %>% 
  filter(huc12_name != "Long Island Sound Deep")



  
plots <- prediction_natural %>% 
  select(44:55) %>% 
  rename("alewife floater" = "alewifefloater", 
         "eastern pearlshell" = "easternpearlshell", 
         "eastern pondmussel" = "easternpondmussel",
         "tidewater mucket" = "Tidewatermucket",
         "yellow lampmussel" = "yellowlampmussel")

for (i in (1:length(names(plots)))) { 

ggplot(data = plots)+
  geom_sf(aes(fill = plots[[i]]), color = NA)+
  scale_fill_gradient2(low = "grey60",
                        mid = "yellow",
                        high = "green3",
                        midpoint = 0.5,
                       limits = c(0,1))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        strip.background = element_rect(fill = "grey50",
                                        colour = "black"),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 15))+
    ggtitle(label = names(plots)[i])+
  labs(fill = "Probability")+
  geom_sf(data =shp[shp$common_name == names(plots)[i],], 
          aes(color = live_occurrence), color = "black", size = .5)

ggsave(paste("tmpfigures/baseline_mussel_predictions/", names(plots)[i], "_natural.png", sep = ""), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300) 

print(i)
}

```







MANAGEMENT AND CLIMATE CHANGE: huc12_damden_sqkm = 0
 
```{r}

#load the 2080 file that has the projected stream temperature and streamflow covariate values
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates_2080.RData")


#need to get the variables named the same as the baseline dataframe and need to order the variables in the same way so that we can scale them
test <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  select(7:8, 10:42) %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(1:2, 30:34, 3:27, 35, 28:29)


#management scenario -decrease huc12_damden_sqkm to 0

test$huc12_damden_sqkm <- 0


#load means and sd of the baseline covariates so we can scale the future data appropriates
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_mean.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_sd.RData")



#scale the variables using the means and sd used to scale the variables in the baseline condition
test <- sweep(test, 2, means, FUN = '-') 
test <- sweep(test, 2, sd, FUN = '/')  

#join the scaled variables back to 
NHDplusV2_NewEngCrop_mg_cnt <- cbind((NHDplusV2_NewEngCrop_2080)[,c(1:6, 9)], test)



#Host fish variable 1 for the mussel model  - to be included in the model for all species of mussels:
#The predicted probability of occurrence of all fish (summed by stream reach – representing the total probability of occurrence of all species within a given reach) and averaged over a HUC12 – ultimately representing the average summed probability of occurrence of any fish species within a HUC12
#using the zprob from the prop abund model

file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions_mg_cnt <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop_mg_cnt, type = "zprob")))) #do 1 - predict because we want prob of occurrence, not the prob of a zero
names(predictions_mg_cnt)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop_mg_cnt, type = "zprob"))))
  names(tmp)[1] <- file.list[i]
  
  predictions_mg_cnt <- cbind(predictions_mg_cnt, tmp)
  
  
}

colnames(predictions_mg_cnt) <- gsub("_Mgmt_Paper_propabun_", "pres_", colnames(predictions_mg_cnt)) #the probabun is the title of the model, but we used the zero predict option so we need to change the column names


#join the predictions_mg_cnt to the list of comids
tmp <- cbind(NHDplusV2_NewEngCrop_mg_cnt, predictions_mg_cnt)


#first average by HUC12 and by species and save cc fish predictions
fish_mg_cnt_proboccurrence_huc12 <- tmp %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, 43:95) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(3:55, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, common_name) %>% 
  summarise(mean_prob_occ = mean(prob_occ)) %>% 
  ungroup() %>% 
  mutate(scenario = "mg_cnt", taxa = "fish")
save(fish_mg_cnt_proboccurrence_huc12, file = 'ProjectResults/fish_mg_cnt_proboccurrence.RData')


#average by HUC12 becuase that is the scale we are modeling mussels
fish_host_prob_occ_mg_cnt <- tmp %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, COMID, 43:95) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(4:56, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, COMID) %>% 
  summarise(total_prob_occ = sum(prob_occ)) %>% 
  ungroup() %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(mean_prob_occ = mean(total_prob_occ))
  






#fish host covariate with proportional abundance for the perferred host fish. (Variable 3 and 4)


file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions_mg_cnt <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop_mg_cnt, type = "response"))
names(predictions_mg_cnt)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop_mg_cnt, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions_mg_cnt <- cbind(predictions_mg_cnt, tmp)
  
  
}

#select only the columns for preferred host fish
predictions_mg_cnt <- predictions_mg_cnt %>% 
  select("_Mgmt_Paper_propabun_brown trout.rds", "_Mgmt_Paper_propabun_rainbow trout.rds", "_Mgmt_Paper_propabun_atlantic salmon.rds", "_Mgmt_Paper_propabun_brook trout.rds",
         "_Mgmt_Paper_propabun_yellow perch.rds", "_Mgmt_Paper_propabun_white perch.rds", "_Mgmt_Paper_propabun_largemouth bass.rds", "_Mgmt_Paper_propabun_bluegill.rds", "_Mgmt_Paper_propabun_pumpkinseed.rds", 
         "_Mgmt_Paper_propabun_redbreast sunfish.rds", )
names(predictions_mg_cnt) <- gsub("_Mgmt_Paper_propabun_", "", names(predictions_mg_cnt))
names(predictions_mg_cnt) <- gsub(".rds", "", names(predictions_mg_cnt))
#names(predictions_mg_cnt) <- gsub("\\.", " ", names(predictions_mg_cnt))


#join the predictions to the list of comids
tmp <- cbind(NHDplusV2_NewEngCrop_mg_cnt, predictions_mg_cnt)

#average by HUC10 becuase that is the scale we are modeling mussels
fish_host_propabun_mg_cnt <- tmp %>%
  data.frame() %>% 
  select(COMID, huc12_name, huc12_tnmid, 43:52) %>% 
  filter(!is.na(brook.trout),
         !is.na(huc12_name)) %>%
  mutate(sumYLamp = white.perch + yellow.perch,
         sumEPear = brown.trout + rainbow.trout + atlantic.salmon + brook.trout,
         sumEpond = yellow.perch + largemouth.bass + bluegill + pumpkinseed + redbreast.sunfish) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(TMuck_perferredhost = mean(white.perch),
            YLamp_perferredhost = mean(sumYLamp),
            EPear_perferredhost = mean(sumEPear),
            Epond_perferredhost = mean(sumEpond))




#Variable at the HUC12 scale showing if the non-modeled host fish has occurred in that HUC12 (Variable 2)
#fish data
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fish_event_huc_join.RData")
load('C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_presence.RData')

fish_event_huc_join <- fish_event_huc_join %>% 
  data.frame() %>% 
  select(-geometry)

df <- left_join(fish_presence, fish_event_huc_join, by = "UID") %>% 
  filter(stock == "natural") %>% 
  select(huc12_name, huc12_tnmid, common_name) %>% 
  unique() %>% 
  filter(common_name %in% c("striped bass", 
                            "rainbow smelt", "american shad", "alewife", "blueback herring",
                            "ninespine stickleback", "three-spined stickleback", "fourspine stickleback"))


huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  data.frame() %>% 
  select(name, tnmid) %>%
  rename(huc12_name = name, huc12_tnmid = tnmid) %>% 
  unique()
fish <- unique(df$common_name)

#make a dataframe that has each UID repeated for each fish spp
absence <- data.frame('huc12_name' = rep(huc12$huc12_name, each = 8,),
                      'huc12_tnmid' = rep(huc12$huc12_tnmid, each = 8))
absence$common_name <- rep(fish, times = nrow(huc12))
#keep only rows in the absence list that are no in the fish_presence list
keep <- anti_join(absence, df, by = c("huc12_name", "huc12_tnmid", "common_name"))

#add presence absence information
keep$occurrence <- 0
df$occurrence <- 1



#bind the presence and absence columns together
fish_host_occ <- rbind(df, keep)
fish_host_occ <- fish_host_occ %>% 
  arrange(huc12_name, huc12_tnmid) %>% 
  pivot_wider(names_from = common_name, values_from = occurrence)

rm(absence, fish_presence, keep)
rm(fish, huc12)

fish_host_occ$YLamp_notmodeledhost <- fish_host_occ$`striped bass`
fish_host_occ$Epond_notmodeledhost <- ifelse(fish_host_occ$`three-spined stickleback` == 1 |
                                                 fish_host_occ$`fourspine stickleback` == 1 |
                                                 fish_host_occ$`ninespine stickleback` == 1,
                                               1, 0)
fish_host_occ$Afloa_notmodeledhost <- ifelse(fish_host_occ$alewife == 1 |
                                                 fish_host_occ$`blueback herring` == 1 |
                                                 fish_host_occ$`american shad` == 1 |
                                                 fish_host_occ$`rainbow smelt` == 1,
                                               1, 0)


fish_host_occ <- fish_host_occ %>% 
  select(huc12_name, huc12_tnmid, YLamp_notmodeledhost, Epond_notmodeledhost, Afloa_notmodeledhost)






```







apply the predictions to the climate change  projections for flow AND temperature combined with the management for dam density


```{r}

load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_mussel_covariates_HUC12.RData")

covariates_mg_cnt <- left_join(NHDplusV2_NewEngCrop_mussel_covariates_HUC12, fish_host_prob_occ_mg_cnt, by = c("huc12_name", "huc12_tnmid"))

covariates_mg_cnt <- left_join(covariates_mg_cnt, fish_host_propabun_mg_cnt, by = c("huc12_name", "huc12_tnmid"))

covariates_mg_cnt <- left_join(covariates_mg_cnt, fish_host_occ, by = c("huc12_name", "huc12_tnmid"))


#load the 2080 file that has the projected stream temperature and streamflow covariate values
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates_2080.RData")


#average by huc12 all the projected flow and temperature variables
test <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(huc12_name, huc12_tnmid, annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(annual_mean_summer_temp = mean(annual_mean_summer_temp, na.rm = T),
            BFI_HIST = mean(BFI_HIST, na.rm = T),
            LO7Q1DT_HIST = mean(LO7Q1DT_HIST, na.rm = T),
            CFM_HIST = mean(CFM_HIST, na.rm = T),
            W95_HIST = mean(W95_HIST, na.rm = T),
            logMJJA_HIST = mean(logMJJA_HIST, na.rm = T))


#need to order the new dataframe correctly and scale using the means and sd from the baseline data

#remove the current flow and temperature values from the baseline file and add in the 2080 projections
order <- names(covariates_mg_cnt)
covariates_mg_cnt2 <- covariates_mg_cnt %>% 
  select(-c(annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST)) %>% #remove the baseline values for these variables
  left_join(test, by = c("huc12_name", "huc12_tnmid")) %>%  #add in the projected 2080 values
  select(all_of(order))

#management scenario #1 - water quality scenario (decrease huc12_damden_sqkm to 0%)

covariates_mg_cnt2$huc12_damden_sqkm <- 0

#load the values used to scale the mussel covariates
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_mean.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_sd.RData")



#scale the variables using the means and sd used to scale the variables in the baseline condition
test <- covariates_mg_cnt2[,c(3:42)]
test <- sweep(test, 2, means, FUN = '-') 
test <- sweep(test, 2, sd, FUN = '/')  

#join the scaled variables back to the full dataset
covariates_mg_cnt <- cbind((covariates_mg_cnt2)[,c(1:2)], test)



#apply models to the climate change and management covariates
file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "mussel_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = predict(mdl, newdata = covariates_mg_cnt, type = "response"))
names(predictions)[1] <- file.list[i]




for (i in 2:12) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = covariates_mg_cnt, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  
  
}

names(predictions) <- gsub(pattern = ".rds", replacement = "", x = names(predictions))
names(predictions) <- gsub(pattern = "mussel_", replacement = "", x = names(predictions))

#join the predictions to the list of comids
prediction_mg_cnt <- cbind(covariates_mg_cnt, predictions)

huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  select(tnmid, areasqkm , name)%>% 
  rename(huc12_areasqkm = areasqkm, huc12_name= name, huc12_tnmid = tnmid)
huc12 <- huc12[!duplicated(huc12$huc12_tnmid),]#Remove the duplicated TNMID fields

prediction_mg_cnt <- left_join(huc12, prediction_mg_cnt, by = c("huc12_name", "huc12_tnmid"))


#save data files with mussel predictions for climate change/landcover mgmt
save(prediction_mg_cnt, file = 'ProjectResults/mussel_CC2080_connectivity_mgmt_proboccurrence.RData')


#sample plot
shp <- left_join(mussel_event_huc_join, mussel_occurrence, by = "UID") %>% 
  filter(!is.na(live_occurrence)) %>% 
  filter(live_occurrence == 1)



  
plots <- prediction_mg_cnt %>% 
  select(44:55) %>% 
  rename("alewife floater" = "alewifefloater", 
         "eastern pearlshell" = "easternpearlshell", 
         "eastern pondmussel" = "easternpondmussel",
         "tidewater mucket" = "Tidewatermucket",
         "yellow lampmussel" = "yellowlampmussel")

for (i in (1:length(names(plots)))) { 

ggplot(data = plots)+
  geom_sf(aes(fill = plots[[i]]), color = NA)+
  scale_fill_gradient2(low = "grey60",
                        mid = "yellow",
                        high = "green3",
                        midpoint = 0.5,
                       limits = c(0,1))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        strip.background = element_rect(fill = "grey50",
                                        colour = "black"),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 15))+
    ggtitle(label = names(plots)[i])+
  labs(fill = "Probability")+
  geom_sf(data =shp[shp$common_name == names(plots)[i],], 
          aes(color = live_occurrence), color = "black", size = .5)

ggsave(paste("tmpfigures/baseline_mussel_predictions/", names(plots)[i], "_mgmt_cnt.png", sep = ""), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300) 

print(i)
}

```















MANAGEMENT AND CLIMATE CHANGE: pctForest_ws = 100 
 Now, calculate mussel probability of occurrence with the climate change AND management scenario for 2080
 first, re-calculate the host fish covariates with the climate change data and improved forest
```{r}

#load the 2080 file that has the projected stream temperature and streamflow covariate values
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates_2080.RData")


#need to get the variables named the same as the baseline dataframe and need to order the variables in the same way so that we can scale them
test <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  select(7:8, 10:42) %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(1:2, 30:34, 3:27, 35, 28:29)


#management - pctForest_ws = 100

test$pctForest_ws <- 100


#load means and sd of the baseline covariates so we can scale the future data appropriates
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_mean.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_sd.RData")



#scale the variables using the means and sd used to scale the variables in the baseline condition
test <- sweep(test, 2, means, FUN = '-') 
test <- sweep(test, 2, sd, FUN = '/')  

#join the scaled variables back to 
NHDplusV2_NewEngCrop_forest <- cbind((NHDplusV2_NewEngCrop_2080)[,c(1:6, 9)], test)



#Host fish variable 1 for the mussel model  - to be included in the model for all species of mussels:
#The predicted probability of occurrence of all fish (summed by stream reach – representing the total probability of occurrence of all species within a given reach) and averaged over a HUC12 – ultimately representing the average summed probability of occurrence of any fish species within a HUC12
#using the zprob from the prop abund model

file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions_forest <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop_forest, type = "zprob")))) #do 1 - predict because we want prob of occurrence, not the prob of a zero
names(predictions_forest)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop_forest, type = "zprob"))))
  names(tmp)[1] <- file.list[i]
  
  predictions_forest <- cbind(predictions_forest, tmp)
  
  
}

colnames(predictions_forest) <- gsub("_Mgmt_Paper_propabun_", "pres_", colnames(predictions_forest)) #the probabun is the title of the model, but we used the zero predict option so we need to change the column names


#join the predictions_forest to the list of comids
tmp <- cbind(NHDplusV2_NewEngCrop_forest, predictions_forest)


#first average by HUC12 and by species and save cc fish predictions
fish_forest_proboccurrence_huc12 <- tmp %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, 43:95) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(3:55, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, common_name) %>% 
  summarise(mean_prob_occ = mean(prob_occ)) %>% 
  ungroup() %>% 
  mutate(scenario = "forest", taxa = "fish")
save(fish_forest_proboccurrence_huc12, file = 'ProjectResults/fish_forest_proboccurrence.RData')


#average by HUC12 becuase that is the scale we are modeling mussels
fish_host_prob_occ_forest <- tmp %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, COMID, 43:95) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(4:56, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, COMID) %>% 
  summarise(total_prob_occ = sum(prob_occ)) %>% 
  ungroup() %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(mean_prob_occ = mean(total_prob_occ))
  






#fish host covariate with proportional abundance for the perferred host fish. (Variable 3 and 4)


file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions_forest <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop_forest, type = "response"))
names(predictions_forest)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop_forest, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions_forest <- cbind(predictions_forest, tmp)
  
  
}

#select only the columns for preferred host fish
predictions_forest <- predictions_forest %>% 
  select("_Mgmt_Paper_propabun_brown trout.rds", "_Mgmt_Paper_propabun_rainbow trout.rds", "_Mgmt_Paper_propabun_atlantic salmon.rds", "_Mgmt_Paper_propabun_brook trout.rds",
         "_Mgmt_Paper_propabun_yellow perch.rds", "_Mgmt_Paper_propabun_white perch.rds", "_Mgmt_Paper_propabun_largemouth bass.rds", "_Mgmt_Paper_propabun_bluegill.rds", "_Mgmt_Paper_propabun_pumpkinseed.rds", 
         "_Mgmt_Paper_propabun_redbreast sunfish.rds", )
names(predictions_forest) <- gsub("_Mgmt_Paper_propabun_", "", names(predictions_forest))
names(predictions_forest) <- gsub(".rds", "", names(predictions_forest))
#names(predictions_forest) <- gsub("\\.", " ", names(predictions_forest))


#join the predictions to the list of comids
tmp <- cbind(NHDplusV2_NewEngCrop_forest, predictions_forest)

#average by HUC10 becuase that is the scale we are modeling mussels
fish_host_propabun_forest <- tmp %>%
  data.frame() %>% 
  select(COMID, huc12_name, huc12_tnmid, 43:52) %>% 
  filter(!is.na(brook.trout),
         !is.na(huc12_name)) %>%
  mutate(sumYLamp = white.perch + yellow.perch,
         sumEPear = brown.trout + rainbow.trout + atlantic.salmon + brook.trout,
         sumEpond = yellow.perch + largemouth.bass + bluegill + pumpkinseed + redbreast.sunfish) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(TMuck_perferredhost = mean(white.perch),
            YLamp_perferredhost = mean(sumYLamp),
            EPear_perferredhost = mean(sumEPear),
            Epond_perferredhost = mean(sumEpond))




#Variable at the HUC12 scale showing if the non-modeled host fish has occurred in that HUC12 (Variable 2)
#fish data
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fish_event_huc_join.RData")
load('C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_presence.RData')

fish_event_huc_join <- fish_event_huc_join %>% 
  data.frame() %>% 
  select(-geometry)

df <- left_join(fish_presence, fish_event_huc_join, by = "UID") %>% 
  filter(stock == "natural") %>% 
  select(huc12_name, huc12_tnmid, common_name) %>% 
  unique() %>% 
  filter(common_name %in% c("striped bass", 
                            "rainbow smelt", "american shad", "alewife", "blueback herring",
                            "ninespine stickleback", "three-spined stickleback", "fourspine stickleback"))


huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  data.frame() %>% 
  select(name, tnmid) %>%
  rename(huc12_name = name, huc12_tnmid = tnmid) %>% 
  unique()
fish <- unique(df$common_name)

#make a dataframe that has each UID repeated for each fish spp
absence <- data.frame('huc12_name' = rep(huc12$huc12_name, each = 8,),
                      'huc12_tnmid' = rep(huc12$huc12_tnmid, each = 8))
absence$common_name <- rep(fish, times = nrow(huc12))
#keep only rows in the absence list that are no in the fish_presence list
keep <- anti_join(absence, df, by = c("huc12_name", "huc12_tnmid", "common_name"))

#add presence absence information
keep$occurrence <- 0
df$occurrence <- 1



#bind the presence and absence columns together
fish_host_occ <- rbind(df, keep)
fish_host_occ <- fish_host_occ %>% 
  arrange(huc12_name, huc12_tnmid) %>% 
  pivot_wider(names_from = common_name, values_from = occurrence)

rm(absence, fish_presence, keep)
rm(fish, huc12)

fish_host_occ$YLamp_notmodeledhost <- fish_host_occ$`striped bass`
fish_host_occ$Epond_notmodeledhost <- ifelse(fish_host_occ$`three-spined stickleback` == 1 |
                                                 fish_host_occ$`fourspine stickleback` == 1 |
                                                 fish_host_occ$`ninespine stickleback` == 1,
                                               1, 0)
fish_host_occ$Afloa_notmodeledhost <- ifelse(fish_host_occ$alewife == 1 |
                                                 fish_host_occ$`blueback herring` == 1 |
                                                 fish_host_occ$`american shad` == 1 |
                                                 fish_host_occ$`rainbow smelt` == 1,
                                               1, 0)


fish_host_occ <- fish_host_occ %>% 
  select(huc12_name, huc12_tnmid, YLamp_notmodeledhost, Epond_notmodeledhost, Afloa_notmodeledhost)






```







apply the predictions to the climate change  projections for flow AND temperature combined with the management for forest


```{r}

load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_mussel_covariates_HUC12.RData")

covariates_forest <- left_join(NHDplusV2_NewEngCrop_mussel_covariates_HUC12, fish_host_prob_occ_forest, by = c("huc12_name", "huc12_tnmid"))

covariates_forest <- left_join(covariates_forest, fish_host_propabun_forest, by = c("huc12_name", "huc12_tnmid"))

covariates_forest <- left_join(covariates_forest, fish_host_occ, by = c("huc12_name", "huc12_tnmid"))


#load the 2080 file that has the projected stream temperature and streamflow covariate values
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates_2080.RData")


#average by huc12 all the projected flow and temperature variables
test <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(huc12_name, huc12_tnmid, annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(annual_mean_summer_temp = mean(annual_mean_summer_temp, na.rm = T),
            BFI_HIST = mean(BFI_HIST, na.rm = T),
            LO7Q1DT_HIST = mean(LO7Q1DT_HIST, na.rm = T),
            CFM_HIST = mean(CFM_HIST, na.rm = T),
            W95_HIST = mean(W95_HIST, na.rm = T),
            logMJJA_HIST = mean(logMJJA_HIST, na.rm = T))


#need to order the new dataframe correctly and scale using the means and sd from the baseline data

#remove the current flow and temperature values from the baseline file and add in the 2080 projections
order <- names(covariates_forest)
covariates_forest2 <- covariates_forest %>% 
  select(-c(annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST)) %>% #remove the baseline values for these variables
  left_join(test, by = c("huc12_name", "huc12_tnmid")) %>%  #add in the projected 2080 values
  select(all_of(order))

#management scenario  -  increase forest to 100%

covariates_forest2$pctForest_ws <- 100

#load the values used to scale the mussel covariates
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_mean.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_sd.RData")



#scale the variables using the means and sd used to scale the variables in the baseline condition
test <- covariates_forest2[,c(3:42)]
test <- sweep(test, 2, means, FUN = '-') 
test <- sweep(test, 2, sd, FUN = '/')  

#join the scaled variables back to the full dataset
covariates_forest <- cbind((covariates_forest2)[,c(1:2)], test)



#apply models to the climate change and management covariates
file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "mussel_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = predict(mdl, newdata = covariates_forest, type = "response"))
names(predictions)[1] <- file.list[i]




for (i in 2:12) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = covariates_forest, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  
  
}

names(predictions) <- gsub(pattern = ".rds", replacement = "", x = names(predictions))
names(predictions) <- gsub(pattern = "mussel_", replacement = "", x = names(predictions))

#join the predictions to the list of comids
prediction_forest <- cbind(covariates_forest, predictions)

huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  select(tnmid, areasqkm , name)%>% 
  rename(huc12_areasqkm = areasqkm, huc12_name= name, huc12_tnmid = tnmid)
huc12 <- huc12[!duplicated(huc12$huc12_tnmid),]#Remove the duplicated TNMID fields

prediction_forest <- left_join(huc12, prediction_forest, by = c("huc12_name", "huc12_tnmid"))


#save data files with mussel predictions for climate change/landcover mgmt
save(prediction_forest, file = 'ProjectResults/mussel_CC2080_forest_proboccurrence.RData')


#sample plot
shp <- left_join(mussel_event_huc_join, mussel_occurrence, by = "UID") %>% 
  filter(!is.na(live_occurrence)) %>% 
  filter(live_occurrence == 1)



  
plots <- prediction_forest %>% 
  select(44:55) %>% 
  rename("alewife floater" = "alewifefloater", 
         "eastern pearlshell" = "easternpearlshell", 
         "eastern pondmussel" = "easternpondmussel",
         "tidewater mucket" = "Tidewatermucket",
         "yellow lampmussel" = "yellowlampmussel")

for (i in (1:length(names(plots)))) { 

ggplot(data = plots)+
  geom_sf(aes(fill = plots[[i]]), color = NA)+
  scale_fill_gradient2(low = "grey60",
                        mid = "yellow",
                        high = "green3",
                        midpoint = 0.5,
                       limits = c(0,1))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        strip.background = element_rect(fill = "grey50",
                                        colour = "black"),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 15))+
    ggtitle(label = names(plots)[i])+
  labs(fill = "Probability")+
  geom_sf(data =shp[shp$common_name == names(plots)[i],], 
          aes(color = live_occurrence), color = "black", size = .5)

ggsave(paste("tmpfigures/baseline_mussel_predictions/", names(plots)[i], "_forest.png", sep = ""), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300) 

print(i)
}

```




MANAGEMENT and climate change- combine all
PctImp_WsRp100 = 0, 
pctNatural_ripbuf100 = 100,
forest = 100
dam density = 0 

calculate mussel probability of occurrence with the climate change AND management scenario for 2080
first, re-calculate the host fish covariates with the climate change data and mgmt scenario 1 
```{r}

#load the 2080 file that has the projected stream temperature and streamflow covariate values
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates_2080.RData")


#need to get the variables named the same as the baseline dataframe and need to order the variables in the same way so that we can scale them
test <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  select(7:8, 10:42) %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(1:2, 30:34, 3:27, 35, 28:29)


#management scenario - combine all management options

test$PctImp_WsRp100 <- 0
test$pctNatural_ripbuf100 <- 100
test$pctForest_ws <- 100
test$huc12_damden_sqkm <- 0


#load means and sd of the baseline covariates so we can scale the future data appropriates
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_mean.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fishcovariates_scaling_sd.RData")



#scale the variables using the means and sd used to scale the variables in the baseline condition
test <- sweep(test, 2, means, FUN = '-') 
test <- sweep(test, 2, sd, FUN = '/')  

#join the scaled variables back to 
NHDplusV2_NewEngCrop_mg <- cbind((NHDplusV2_NewEngCrop_2080)[,c(1:6, 9)], test)



#Host fish variable 1 for the mussel model  - to be included in the model for all species of mussels:
#The predicted probability of occurrence of all fish (summed by stream reach – representing the total probability of occurrence of all species within a given reach) and averaged over a HUC12 – ultimately representing the average summed probability of occurrence of any fish species within a HUC12
#using the zprob from the prop abund model

file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions_mg <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop_mg, type = "zprob")))) #do 1 - predict because we want prob of occurrence, not the prob of a zero
names(predictions_mg)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = (1 - (predict(mdl, newdata = NHDplusV2_NewEngCrop_mg, type = "zprob"))))
  names(tmp)[1] <- file.list[i]
  
  predictions_mg <- cbind(predictions_mg, tmp)
  
  
}

colnames(predictions_mg) <- gsub("_Mgmt_Paper_propabun_", "pres_", colnames(predictions_mg)) #the probabun is the title of the model, but we used the zero predict option so we need to change the column names


#join the predictions_mg to the list of comids
tmp <- cbind(NHDplusV2_NewEngCrop_mg, predictions_mg)


#first average by HUC12 and by species and save cc fish predictions
fish_mg_proboccurrence_huc12 <- tmp %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, 43:95) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(3:55, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, common_name) %>% 
  summarise(mean_prob_occ = mean(prob_occ)) %>% 
  ungroup() %>% 
  mutate(scenario = "mg", taxa = "fish")
save(fish_mg_proboccurrence_huc12, file = 'ProjectResults/fish_mg_proboccurrence.RData')


#average by HUC12 becuase that is the scale we are modeling mussels
fish_host_prob_occ_mg <- tmp %>%
  data.frame() %>% 
  select(huc12_name, huc12_tnmid, COMID, 43:95) %>% 
  filter(!is.na(pres_brook.trout.rds),
         !is.na(huc12_name)) %>%
  pivot_longer(4:56, names_to = "common_name", values_to = "prob_occ") %>% 
  group_by(huc12_name, huc12_tnmid, COMID) %>% 
  summarise(total_prob_occ = sum(prob_occ)) %>% 
  ungroup() %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(mean_prob_occ = mean(total_prob_occ))
  






#fish host covariate with proportional abundance for the perferred host fish. (Variable 3 and 4)


file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "_Mgmt_Paper_propabun_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions_mg <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop_mg, type = "response"))
names(predictions_mg)[1] <- file.list[i]




for (i in 2:53) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = NHDplusV2_NewEngCrop_mg, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions_mg <- cbind(predictions_mg, tmp)
  
  
}

#select only the columns for preferred host fish
predictions_mg <- predictions_mg %>% 
  select("_Mgmt_Paper_propabun_brown trout.rds", "_Mgmt_Paper_propabun_rainbow trout.rds", "_Mgmt_Paper_propabun_atlantic salmon.rds", "_Mgmt_Paper_propabun_brook trout.rds",
         "_Mgmt_Paper_propabun_yellow perch.rds", "_Mgmt_Paper_propabun_white perch.rds", "_Mgmt_Paper_propabun_largemouth bass.rds", "_Mgmt_Paper_propabun_bluegill.rds", "_Mgmt_Paper_propabun_pumpkinseed.rds", 
         "_Mgmt_Paper_propabun_redbreast sunfish.rds", )
names(predictions_mg) <- gsub("_Mgmt_Paper_propabun_", "", names(predictions_mg))
names(predictions_mg) <- gsub(".rds", "", names(predictions_mg))
#names(predictions_mg) <- gsub("\\.", " ", names(predictions_mg))


#join the predictions to the list of comids
tmp <- cbind(NHDplusV2_NewEngCrop_mg, predictions_mg)

#average by HUC10 becuase that is the scale we are modeling mussels
fish_host_propabun_mg <- tmp %>%
  data.frame() %>% 
  select(COMID, huc12_name, huc12_tnmid, 43:52) %>% 
  filter(!is.na(brook.trout),
         !is.na(huc12_name)) %>%
  mutate(sumYLamp = white.perch + yellow.perch,
         sumEPear = brown.trout + rainbow.trout + atlantic.salmon + brook.trout,
         sumEpond = yellow.perch + largemouth.bass + bluegill + pumpkinseed + redbreast.sunfish) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(TMuck_perferredhost = mean(white.perch),
            YLamp_perferredhost = mean(sumYLamp),
            EPear_perferredhost = mean(sumEPear),
            Epond_perferredhost = mean(sumEpond))




#Variable at the HUC12 scale showing if the non-modeled host fish has occurred in that HUC12 (Variable 2)
#fish data
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fish_event_huc_join.RData")
load('C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_presence.RData')

fish_event_huc_join <- fish_event_huc_join %>% 
  data.frame() %>% 
  select(-geometry)

df <- left_join(fish_presence, fish_event_huc_join, by = "UID") %>% 
  filter(stock == "natural") %>% 
  select(huc12_name, huc12_tnmid, common_name) %>% 
  unique() %>% 
  filter(common_name %in% c("striped bass", 
                            "rainbow smelt", "american shad", "alewife", "blueback herring",
                            "ninespine stickleback", "three-spined stickleback", "fourspine stickleback"))


huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  data.frame() %>% 
  select(name, tnmid) %>%
  rename(huc12_name = name, huc12_tnmid = tnmid) %>% 
  unique()
fish <- unique(df$common_name)

#make a dataframe that has each UID repeated for each fish spp
absence <- data.frame('huc12_name' = rep(huc12$huc12_name, each = 8,),
                      'huc12_tnmid' = rep(huc12$huc12_tnmid, each = 8))
absence$common_name <- rep(fish, times = nrow(huc12))
#keep only rows in the absence list that are no in the fish_presence list
keep <- anti_join(absence, df, by = c("huc12_name", "huc12_tnmid", "common_name"))

#add presence absence information
keep$occurrence <- 0
df$occurrence <- 1



#bind the presence and absence columns together
fish_host_occ <- rbind(df, keep)
fish_host_occ <- fish_host_occ %>% 
  arrange(huc12_name, huc12_tnmid) %>% 
  pivot_wider(names_from = common_name, values_from = occurrence)

rm(absence, fish_presence, keep)
rm(fish, huc12)

fish_host_occ$YLamp_notmodeledhost <- fish_host_occ$`striped bass`
fish_host_occ$Epond_notmodeledhost <- ifelse(fish_host_occ$`three-spined stickleback` == 1 |
                                                 fish_host_occ$`fourspine stickleback` == 1 |
                                                 fish_host_occ$`ninespine stickleback` == 1,
                                               1, 0)
fish_host_occ$Afloa_notmodeledhost <- ifelse(fish_host_occ$alewife == 1 |
                                                 fish_host_occ$`blueback herring` == 1 |
                                                 fish_host_occ$`american shad` == 1 |
                                                 fish_host_occ$`rainbow smelt` == 1,
                                               1, 0)


fish_host_occ <- fish_host_occ %>% 
  select(huc12_name, huc12_tnmid, YLamp_notmodeledhost, Epond_notmodeledhost, Afloa_notmodeledhost)




#the correct host fish files for the management scenario include: fish_host_occ, fish_host_prob_occ_cc, fish_host_propabun_cc

```







apply the predictions to the climate change  projections for flow AND temperature combined with the management for Imperviousnes and natural riparian buffer


```{r}

load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_mussel_covariates_HUC12.RData")

covariates_mg <- left_join(NHDplusV2_NewEngCrop_mussel_covariates_HUC12, fish_host_prob_occ_mg, by = c("huc12_name", "huc12_tnmid"))

covariates_mg <- left_join(covariates_mg, fish_host_propabun_mg, by = c("huc12_name", "huc12_tnmid"))

covariates_mg <- left_join(covariates_mg, fish_host_occ, by = c("huc12_name", "huc12_tnmid"))


#load the 2080 file that has the projected stream temperature and streamflow covariate values
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/NHDplusV2_NewEngCrop_covariates_2080.RData")


#average by huc12 all the projected flow and temperature variables
test <- NHDplusV2_NewEngCrop_2080 %>% 
  data.frame() %>% 
  rename(annual_mean_summer_temp = annual_mean_summer_temp_pls_4,
         BFI_HIST = BFI_2080,
         LO7Q1DT_HIST = LO7Q1DT_2080,
         CFM_HIST = CFM_2080,
         W95_HIST = W95_2080,
         logMJJA_HIST = logMJJA_2080) %>% 
  select(huc12_name, huc12_tnmid, annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST) %>% 
  group_by(huc12_name, huc12_tnmid) %>% 
  summarise(annual_mean_summer_temp = mean(annual_mean_summer_temp, na.rm = T),
            BFI_HIST = mean(BFI_HIST, na.rm = T),
            LO7Q1DT_HIST = mean(LO7Q1DT_HIST, na.rm = T),
            CFM_HIST = mean(CFM_HIST, na.rm = T),
            W95_HIST = mean(W95_HIST, na.rm = T),
            logMJJA_HIST = mean(logMJJA_HIST, na.rm = T))


#need to order the new dataframe correctly and scale using the means and sd from the baseline data

#remove the current flow and temperature values from the baseline file and add in the 2080 projections
order <- names(covariates_mg)
covariates_mg2 <- covariates_mg %>% 
  select(-c(annual_mean_summer_temp, BFI_HIST, LO7Q1DT_HIST, CFM_HIST, W95_HIST, logMJJA_HIST)) %>% #remove the baseline values for these variables
  left_join(test, by = c("huc12_name", "huc12_tnmid")) %>%  #add in the projected 2080 values
  select(all_of(order))

#management scenario - combine all managemnet

covariates_mg2$PctImp_WsRp100 <- 0
covariates_mg2$pctNatural_ripbuf100  <- 100
covariates_mg2$pctForest_ws  <- 100
covariates_mg2$huc12_damden_sqkm  <- 0


#load the values used to scale the mussel covariates
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_mean.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/musselcovariates_scaling_sd.RData")



#scale the variables using the means and sd used to scale the variables in the baseline condition
test <- covariates_mg2[,c(3:42)]
test <- sweep(test, 2, means, FUN = '-') 
test <- sweep(test, 2, sd, FUN = '/')  

#join the scaled variables back to the full dataset
covariates_mg <- cbind((covariates_mg2)[,c(1:2)], test)



#apply models to the climate change and management covariates
file.list <- list.files("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects", pattern = "mussel_")

i <- 1

mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
predictions <- data.frame("test" = predict(mdl, newdata = covariates_mg, type = "response"))
names(predictions)[1] <- file.list[i]




for (i in 2:12) { 
  
  
   mdl <- readRDS(paste("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_objects/", file.list[i], sep = ""))
  
  tmp <- data.frame("test" = predict(mdl, newdata = covariates_mg, type = "response"))
  names(tmp)[1] <- file.list[i]
  
  predictions <- cbind(predictions, tmp)
  
  
}

names(predictions) <- gsub(pattern = ".rds", replacement = "", x = names(predictions))
names(predictions) <- gsub(pattern = "mussel_", replacement = "", x = names(predictions))

#join the predictions to the list of comids
prediction_mg <- cbind(covariates_mg, predictions)

huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
huc12 <- huc12 %>% 
  select(tnmid, areasqkm , name)%>% 
  rename(huc12_areasqkm = areasqkm, huc12_name= name, huc12_tnmid = tnmid)
huc12 <- huc12[!duplicated(huc12$huc12_tnmid),]#Remove the duplicated TNMID fields

prediction_mg <- left_join(huc12, prediction_mg, by = c("huc12_name", "huc12_tnmid"))


#save data files with mussel predictions for climate change/landcover mgmt
save(prediction_mg, file = 'ProjectResults/mussel_CC2080_landcover_mgmt_proboccurrence.RData')


#sample plot
shp <- left_join(mussel_event_huc_join, mussel_occurrence, by = "UID") %>% 
  filter(!is.na(live_occurrence)) %>% 
  filter(live_occurrence == 1) %>% 
  filter(huc12_name != "Long Island Sound Deep")



  
plots <- prediction_mg %>% 
  select(44:55) %>% 
  rename("alewife floater" = "alewifefloater", 
         "eastern pearlshell" = "easternpearlshell", 
         "eastern pondmussel" = "easternpondmussel",
         "tidewater mucket" = "Tidewatermucket",
         "yellow lampmussel" = "yellowlampmussel")

for (i in (1:length(names(plots)))) { 

ggplot(data = plots)+
  geom_sf(aes(fill = plots[[i]]), color = NA)+
  scale_fill_gradient2(low = "grey60",
                        mid = "yellow",
                        high = "green3",
                        midpoint = 0.5,
                       limits = c(0,1))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
        strip.background = element_rect(fill = "grey50",
                                        colour = "black"),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 15))+
    ggtitle(label = names(plots)[i])+
  labs(fill = "Probability")+
  geom_sf(data =shp[shp$common_name == names(plots)[i],], 
          aes(color = live_occurrence), color = "black", size = .5)

ggsave(paste("tmpfigures/baseline_mussel_predictions/", names(plots)[i], "_mgmt.png", sep = ""), 
    plot = last_plot(),
    width = 18,
    height = 12,
    units = "cm",
    dpi = 300) 

print(i)
}

```

