---
title: "Biodiversity Model"
author: "Jenny Rogers"
date: '2022-09-22'
output: html_document
---


In the 

```{r}
library(tidyverse)
library(sf)
library(lubridate)
library(corrplot)


```


In this first chunk, read in files: 
fish counts with zeros, fish event huc join, flow metrics by survey
```{r}
#fish data
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fish_occurrence.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fish_count_with_zeros.RData")

#fish event and hydrography join
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fish_event_huc_join.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fish_event_flowline_join.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/fish_event_flowlineV2_join.RData")

#USGS gauge data
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/flowmetrics/survey_huc8flowmetrics.RData")

#SHEDs temp data
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/sheds_temp_metrics_huc12.RData")
load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/sheds_temp_metrics_annaul.RData")

#watershed shape files
huc8 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU8/WBDHU8_NE.shp")


#USFS flow metrics
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/USFS flow metrics/flowmet_historical_crop.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/USFS flow metrics/flowmet_endofcen_crop.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/USFS flow metrics/flowmet_midcen_crop.RData")


#StreamCat
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/strmcatByyr_census.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/strmcatByyr_landcov.RData") 
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/strmcatByyr_imp.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/strmcat_byCOMID.RData")


```






first make a proportional abundance column and then join the count data to the event data and to the flow data.
 
```{r}
counts <- fish_count_with_zeros %>% 
  group_by(UID) %>% 
  mutate(totalcount = sum(count)) %>% 
  ungroup() %>% 
  mutate(propabun = ((count / totalcount)*100)) %>% 
  select(UID, count, totalcount, propabun, common_name) %>% 
  filter(common_name == "pumpkinseed")




event <- fish_event_huc_join %>% 
  data.frame() %>% 
  select(UID, state, year, month, huc8_name, huc12_name, huc8_areasqkm, huc12_areasqkm)

flow <- sites %>% 
  select(huc8_name, year, month, 14:22)

dat <- left_join(counts, event, by = "UID")
dat <- left_join(dat, flow, by = c("huc8_name", "year", "month"))



#join the temperature data at the huc12 scale by time period
dat <- dat %>% 
  mutate(timeperiod = ifelse(year <= 1999, "pre", "post")) %>% 
  left_join(metrics12, by = c("huc12_name", "timeperiod"))

#join the temperature data at the huc12 scale by year
names(metrics_annual)[3:7] <- paste("annual", names(metrics_annual)[3:7], sep = "_")
dat <- left_join(dat, metrics_annual, by = c("huc12_name", "year"))



#join the flowlineV2 with the COMID
fish_event_flowlineV2_join <- fish_event_flowlineV2_join %>% 
  mutate(long = unlist(map(fish_event_flowlineV2_join$geometry, 1)),
         lat = unlist(map(fish_event_flowlineV2_join$geometry, 2))) %>% 
  data.frame() %>% 
  select(-geometry)

dat <- dat %>% 
  select(-c(state, year, month)) %>% 
  left_join(fish_event_flowlineV2_join, by = "UID")



#join to the USFS flow metrics
flowmet_historical_crop <- flowmet_historical_crop %>% 
  data.frame() %>% 
  select(COMID, TOTDASQKM, WBAREATYPE, 9:34) %>% 
  mutate(COMID = as.integer(COMID))
dat <- dat %>% 
  left_join(flowmet_historical_crop, by = "COMID")


#join to the streamcat data
strmcatByyr_census <- strmcatByyr_census %>% 
  select(-year)
strmcatByyr_landcov <- strmcatByyr_landcov %>% 
  select(-year)

dat <- dat %>% 
  left_join(strmcat_byCOMID, by = "COMID")
dat <- dat %>% 
  left_join(strmcatByyr_census, by = "COMID")#this is the 2010 data... we might want to do this year in the future..
dat <- dat %>% 
  left_join(strmcatByyr_landcov, by = "COMID")#this is 2001 NLCD data.. we probably will want to download a coupel more, but for now, I just did one year. they also have some that just look at the riparian buffer..
dat <- dat %>% 
  left_join(strmcatByyr_imp, by = c("COMID", "timeperiod"))


#we probably want to add in data for road crossings and riparian area 

#organize dataframe
dat <- dat %>% 
  select("UID", "COMID", "lat", "long",
         "GNIS_ID", "GNIS_NAME", "state", 
         "date", "year", "month", "timeperiod",   
         "count", "totalcount", "propabun", "common_name", 
         "waterbody", "WBAREATYPE", "project", "source", "event_to_flowln_dist_m","REACHCODE", "huc8_name", "huc12_name", 
         "LENGTHKM", "huc8_areasqkm",  "huc12_areasqkm", "WsAreaSqKm", 
         "q_mean_rel", "q_max_rel" ,"low_dur_days", "high_dur_days", "min_month", "monthly_min_flow", 
         "max_month", "monthly_max_flow", "max_temp", "mean_jul_temp", "mean_summer_temp", 
         "mean_max_temp_30d", "mean_n_day_gt_22",  "annual_max_temp", "annual_mean_jul_temp", 
         "annual_mean_summer_temp", "annual_mean_max_temp_30d", "annual_mean_n_day_gt_22", 
         "MA_HIST", "MJAN_HIST",  "MFEB_HIST", "MMAR_HIST", "MAPR_HIST", "MMAY_HIST", 
         "MJUN_HIST", "MJUL_HIST", "MAUG_HIST", "MSEP_HIST", "MOCT_HIST", "MNOV_HIST", "MDEC_HIST", "MDJF_HIST", 
         "MMAM_HIST" , "MJJA_HIST" , "MSON_HIST", "HIQ1_5_HIST", "HIQ10_HIST", "HIQ25_HIST", "LO7Q1_HIST", 
         "LO7Q10_HIST" ,"BFI_HIST", "LO7Q1DT_HIST", "CFM_HIST", "W95_HIST",
         "BFIWs", "DamDensWs", "DamNIDStorWs", "DamNrmStorWs", "ElevWs", "NABD_DensWs", 
         "NABD_NIDStorWs", "NABD_NrmStorWs", "WtDepWs", "HUDen_Ws",
         "PopDen_Ws", "PctOw_Ws", "PctIce_Ws","PctUrbOp_Ws", "PctUrbLo_Ws", "PctUrbMd_Ws", "PctUrbHi_Ws",
         "PctBl_Ws", "PctDecid_Ws", "PctConif_Ws", "PctMxFst_Ws", "PctShrb_Ws", "PctGrs_Ws", "PctHay_Ws", 
         "PctCrop_Ws", "PctWdWet_Ws", "PctHbWet_Ws", "PctImp_Ws", "PctImp_WsRp100")

#we did not include 'TOTDASQKM' from the USGS flow metrics because it was identical and sourced from 'WsAreaSqKm' in StreamCat




```

In this code chunk, we sort through all the variables and remove ones that will not be good predictors
1. zero/low variance across the study area
2. high intercorrelation
3. low predictive power or relatiomship to our outcome variable

We might want to explore combining certain variables together with PCA (for example, all the temperature ones)
```{r}
colSums(is.na(dat))


str(dat)

#variance across study area

nzv <- nearZeroVar(dat[,22:100], saveMetrics = T)

nzv$metric <- row.names(nzv)
remove <- nzv$metric[nzv$zeroVar == TRUE | nzv$nzv == TRUE]


dat2 <- dat %>% 
  select(-remove)

#look for correlation between variables from the same dataset (SHEDs, huc8 streamflow, USGS flow metrics, Streamcat, metadata (lat,long, watershed size, elev, length) )


usgsgague <- dat2 %>% 
  select(28:35) 
usgsgague <- usgsgague[complete.cases(usgsgague),] #remove rows with NA
cor <- cor(usgsgague, method = c("spearman")) #make correlation matrix
corrplot(cor, type = "lower") #plot
#removes correlatioms 0.5 and higher
keep1 <- c("low_dur_days", "high_dur_days", "min_month", "monthly_min_flow", "max_month") 
#removes "q_mean_rel", "q_max_rel", "monthly_max_flow

sheds <- dat2 %>% 
  select(36:44)
sheds <- sheds[complete.cases(sheds),]
cor <- cor(sheds, method = c("spearman"))
corrplot(cor, type = "lower")
#these are all incredible correlated so we will just keep one - annual_mean_max_temp_30d

keep2 <- "annual_mean_max_temp_30d" 
#this removes "max_temp", "mean_jul_temp", "mean_summer_temp", "mean_max_temp_30d", "annual_max_temp",
#              "annual_mean_jul_temp", "annual_mean_summer_temp", "annual_mean_n_day_gt_22"


usfs <- dat2 %>% 
  select(44:70)
usfs <- usfs[complete.cases(usfs),] #remove rows with NA
cor <- cor(usfs, method = c("spearman")) #make correlation matrix
corrplot(cor, type = "lower") #plot
#all of the magnitude variables are very correlated so we will just keep one, Mean JJA, and then we will keep the other non-magnitude variables, which were not correlated at all to anything esle

keep3 <- c("MJJA_HIST","BFI_HIST", "LO7Q1DT_HIST", "CFM_HIST", "W95_HIST") 
#this removes "MA_HIST", "MJAN_HIST","MFEB_HIST", "MMAR_HIST", "MAPR_HIST", "MMAY_HIST", "MJUN_HIST", 
#             "MJUL_HIST","MAUG_HIST", "MSEP_HIST", "MOCT_HIST", "MNOV_HIST", "MDEC_HIST", "MDJF_HIST", 
#             "MMAM_HIST", "MSON_HIST", "HIQ1_5_HIST", "HIQ10_HIST", "HIQ25_HIST", 
#             "LO7Q1_HIST", "LO7Q10_HIST"

strmct <- dat2 %>% 
  select(71:91)
strmct <- strmct[complete.cases(strmct),] #remove rows with NA
cor <- cor(strmct, method = c("spearman")) #make correlation matrix
corrplot(cor, type = "lower") #plot
#the high, med, low, and impervious surface, and housing density, and population are all correlated, but not very very high... we migth want to do a pca with these, though that will make it harder to interpret.
keep4 <-  c("BFIWs", "ElevWs", "WtDepWs", "HUDen_Ws", "PopDen_Ws", "PctOw_Ws", "PctUrbOp_Ws",
           "PctBl_Ws", "PctDecid_Ws", "PctConif_Ws", "PctMxFst_Ws", "PctShrb_Ws", "PctGrs_Ws",
           "PctHay_Ws", "PctWdWet_Ws", "PctHbWet_Ws", "PctImp_Ws", "PctImp_WsRp100")
#removes: "PctUrbLo_Ws", "PctUrbMd_Ws", "PctUrbHi_Ws"


dat3 <- dat2 %>% 
  select(count, propabun, lat, long, state, year, month, source, huc8_name,WsAreaSqKm, keep1, keep2, keep3, keep4) 

#now check for correlation among the retained predictors
final <- dat3 %>% 
  select(3,4, 10:39)
final <- final[complete.cases(final),] #remove rows with NA
cor <- cor(final, method = c("spearman")) #make correlation matrix
corrplot(cor, type = "lower") #plot


```
Here, 
logistic regression plots showing the relatiomship of presence and absence for each variable,
linear regiression plots for proportional abundacne
t test with pvalue table for logistic regression plots

```{r}

dat <- dat %>% 
  mutate(occurrence = ifelse(count>0, 1, 0)) %>% 
  select(91, 3,4,22, 25:90)
ggplot(data = dat, mapping = aes(x = CFM_HIST, y = occurrence))+
  geom_jitter(size = 8, alpha = 0.1)+
  geom_smooth(method = "glm", method.args = list(family = "binomial"))


```
