---
title: "fish survey data exploration"
author: "Jenny Rogers"
date: "3/4/2022"
output: html_document
---


load libraries 
```{r }


#load libraries
library(tidyverse)
library(sf)
library(readxl)

```


Load fish data
``` {r}

#load data
load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_event.RData")
load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_size.RData")
load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_count.RData")
load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_method.RData")



```


First we want to remove the lentic site events from all the datasets

```{r}

event <- event %>% 
  filter(waterbody == "lotic")



#What methods are removed by only looking at lotic
method2 <- method %>% 
  filter(UID %in% event$UID)

tmp <- method %>% 
  filter(! gear %in% method2$gear) %>% 
  select(gear) %>% 
  unique()

#what spp are removed by only looking at lotic

fish_count2 <- fish_count %>% 
  filter(UID %in% event$UID)

tmp <- fish_count %>% 
  filter(! common_name %in% fish_count2$common_name) %>% 
  select(common_name) %>% 
  unique()


rm(tmp, method2, fish_count2)




#make new dfs with only the lotic surveys
fish_count <- fish_count %>% 
  filter(UID %in% event$UID)

fish_size <- fish_size %>% 
  filter(UID %in% event$UID)

method <- method %>% 
  filter(UID %in% event$UID)

```

Then we want to remove the fish spp that we think should be removed based on habitat or being too rare

```{r}

tmp <- read_excel("FishToInclude.xlsx", sheet = 1)
keep <- tmp$common_name[tmp$`project team initial thoughts` == "include"]


#make new dfs with only the fish spp to keep
fish_count <- fish_count %>% 
  filter(common_name %in% keep)

fish_size <- fish_size %>% 
  filter(common_name %in% keep)

#make a new method and event df with only the ones correpsonding the the fish we are keeping
df <- fish_count %>% 
  left_join(event, by= "UID") 

keep <- df$UID


event <- event %>% 
  filter(UID %in% keep)

method <- method %>% 
  filter(UID %in% keep)

```


In this next chunk, we want to make plots that explore the data.
1. survey events by year and state
2. proportion of surveys using which types of gear
3. table of fish counts by gear - do we miss anyone?
4. proportaion of the counts of each spp where lengths were measured (to do life history)
5. proportion of surveys with second pass by state, HUC
6. spp observed in surveys of with a second pass
7. how many surveys occur within the same HUC in the same season? - space for time
8. histograms of reach length and efish duration, by state
9. mean spp diversity and range by differnet sized HUCs .huc on x axis, diversity on y axis, at some point it will increase sharply

``` {r}

#surveys by year and state
table(event$year)
ggplot(data = event, mapping = aes(x = year))+
  geom_bar(aes(fill = state))+
  theme(axis.line = element_line(colour = "black"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "grey90"))+
  labs(x = "Year", y = "Count", fill = "State")+
  geom_vline(xintercept = 2004.5, color = "red", linetype = "dashed")+
  geom_vline(xintercept = 1995.5, color = "red", linetype = "dashed")
ggsave(
    filename = "tmpfigures/loticsurveybyyear.png",
    plot = last_plot(),
    width = 15,
    height = 10,
    units = "cm",
    dpi = 300
  )
  

#number of surveys that are backpack e fish vs other
dat <- fish_count %>%
  filter(!is.na(count)) %>% 
 select(UID) %>% 
  unique()
dat <- fish_size %>% 
  select(UID) %>% 
  unique()

#proportion of surveys using which types of gear
test <- method %>%
  left_join(event, by = "UID") %>% 
  filter(!is.na(gear),
         gear != "water quality")
ggplot(data = test)+
  geom_bar(mapping = aes(x = state, fill = gear),
                         position = "fill")+
  theme(axis.line = element_line(colour = "black"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "grey90"))+
  labs(x = "State", y = "Proportion", fill = "Gear")


#gear by HUC



#table of fish counts by gear - do we miss anyone?
test <- fish_count %>%
  left_join(method, by = "UID") %>% 
  group_by(gear, common_name) %>% 
  summarise(count = sum(count, na.rm = T)) %>% 
  pivot_wider(names_from = gear, values_from = count) 
write.csv(test, "tmpfigures/loticfishcountbygear.csv")

#make figure showing proporation of fish count by gear for the fish where <75% of observations were done by backpack efish

test <- fish_count %>%
  left_join(method, by = "UID") %>% #add gear type
  group_by(gear, common_name) %>% 
  summarise(count = sum(count, na.rm = T)) %>% #get total counts of each spp by gear
  filter(!is.na(gear) & !is.na(common_name))

#make temporary df of total counts for each spp
tmp <- test %>% 
  group_by(common_name) %>% 
  summarize(tot = sum(count, na.rm=T)) #total counts of spp from all gear combined
test2 <- test %>% 
  left_join(tmp, by = "common_name") %>%  #add total counts column
  mutate(prop = count/tot) %>%  #add proportion of count from each gear type out of total
  select(-count, -tot) %>% 
  pivot_wider(names_from = gear, values_from = prop) %>% #wide format 
  filter(efish_backpack < 0.75) %>%  #remove fish whose counts are 75%+ from backpack efish
  pivot_longer(cols = 2:15, names_to = "gear", values_to = "prop") #long format
  
  
ggplot(data = test2)+
  geom_bar(mapping = aes(x = common_name, y = prop, fill = gear),
                         position = "fill", stat = "identity")+
  theme(axis.line = element_line(colour = "black"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "grey90"),
        axis.text.x = element_text(angle = 45, hjust=1, size = 13))+
  labs(x = "Fish", y = "Proportion", fill = "Gear")
ggsave(
    filename = "tmpfigures/test2.png",
    plot = last_plot(),
    width = 26,
    height = 17,
    units = "cm",
    dpi = 300
  )
  


#counts of survey goals (total pick up vs selective pick up)
test <- method %>% 
  group_by(goal) %>% 
  summarise(count = n())





#how many surveys were targeted by spp
test <- method %>%  
  filter(goal == "selective pick-up") %>% 
  group_by(target) %>% 
  summarise(count = n())
write.csv(test, "tmpfigures/table.csv")


#total surveys for each spp that also had the targeted surveys so we can see what percentage the targeted surveys were of the total
test2 <- fish %>% 
  filter(common_name %in% test$target) %>% 
  group_by(common_name) %>% 
  summarize(count = n())



#date range of surveys by state
test <- event %>% #range of surveys by state
  group_by(state) %>% 
  summarise(min = min(date, na.rm = T),
            max = max(date, na.rm = T)) %>% 
  unite(min, max, sep = "-", col = "range")
write.csv(test, "tmpfigures/tmp.csv")


#number of surveys for each number of passes
test <- method %>% 
  group_by(efish_runs) %>% 
  summarise(count = n())
write.csv(test, "tmpfigures/tmp.csv")


#histogram showing proportion of surveys that recorded each species and the spp was measured, vs just counted

tmp <- fish_size %>% #make a df showing number of surveys of each fish that was measured
  select(UID, scientific_name, common_name) %>% 
  unique() %>% 
  group_by(scientific_name, common_name) %>% 
  summarise(count_num = n())

tmp2 <- fish_count %>% #make a df showing number of surveys of each fish that was counted
  select(UID, scientific_name, common_name) %>% 
  unique() %>% 
  group_by(scientific_name, common_name) %>% 
  summarise(count_denom = n())

dat <- left_join(tmp2, tmp, by = c("scientific_name", "common_name")) %>%  #join the two togeter and create column that shows proportion of surveys where each fish was measured, vs the total number of surveys that documented a fish
  mutate(prop = round(count_num/count_denom, 2)) %>% 
  arrange(desc(prop))

write.csv(dat, "tmpfigures/tmp.csv")



# number of fish found in first vs second pass
test <- fish_count %>%
  select(UID, run_num, common_name, scientific_name) %>% 
  group_by(UID, run_num) %>%
  summarise(richness = n()) %>% 
  filter(!is.na(run_num))%>% 
  ungroup() %>% 
  group_by(run_num) %>% 
  summarise(richness_mean = round(mean(richness),2),
            n = n()) 
write.csv(test, "tmpfigures/richnessbyrun.csv")

ggplot(data = test, aes(x = run_num, y = richness_mean)) +
  geom_bar(stat = "identity")
```









Here we look at the data spatially 
We want to download the HUC 8, 10, and 12 shapefiles and merge them togeth
we also want to downlaod all the flowlines for each state and merge them together

``` {r}


#load HUC8 data
MAhuc8 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/MA/Shape/WBDHU8.shp")
VThuc8 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/VT/Shape/WBDHU8.shp")
NHhuc8 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/NH/Shape/WBDHU8.shp")
CThuc8 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/CT/Shape/WBDHU8.shp")
RIhuc8 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/RI/Shape/WBDHU8.shp")

#need to make the names all lowercase
names(MAhuc8) <- tolower(names(MAhuc8))
names(CThuc8) <- tolower(names(CThuc8))
names(NHhuc8) <- tolower(names(NHhuc8))
names(VThuc8) <- tolower(names(VThuc8))
names(RIhuc8) <- tolower(names(RIhuc8))

#rbind HUC8
dat <- bind_rows(MAhuc8, VThuc8, NHhuc8, CThuc8, RIhuc8)

rm(MAhuc8, CThuc8, VThuc8, NHhuc8, RIhuc8)
length(unique(dat$huc8)) #50  - there should be 50 rows

#there are duplicate entries becaseu some watersheds cross state lines
dat8 <- dat[!duplicated(dat$huc8),]
unique(dat8$name)






#load HUC10 data
MAhuc10 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/MA/Shape/WBDHU10.shp")
VThuc10 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/VT/Shape/WBDHU10.shp")
NHhuc10 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/NH/Shape/WBDHU10.shp")
CThuc10 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/CT/Shape/WBDHU10.shp")
RIhuc10 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/RI/Shape/WBDHU10.shp")

#need to make the names all lowercase
names(MAhuc10) <- tolower(names(MAhuc10))
names(CThuc10) <- tolower(names(CThuc10))
names(NHhuc10) <- tolower(names(NHhuc10))
names(VThuc10) <- tolower(names(VThuc10))
names(RIhuc10) <- tolower(names(RIhuc10))

#rbind HUC10
dat <- bind_rows(MAhuc10, VThuc10, NHhuc10, CThuc10, RIhuc10)

rm(MAhuc10, CThuc10, VThuc10, NHhuc10, RIhuc10)
length(unique(dat$huc10)) #50  - there should be 315 rows

#there are duplicate entries becaseu some watersheds cross state lines
dat10 <- dat[!duplicated(dat$huc10),]
unique(dat10$name)







#load HUC12 data
MAhuc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/MA/Shape/WBDHU12.shp")
VThuc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/VT/Shape/WBDHU12.shp")
NHhuc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/NH/Shape/WBDHU12.shp")
CThuc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/CT/Shape/WBDHU12.shp")
RIhuc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/RI/Shape/WBDHU12.shp")

#need to make the names all lowercase
names(MAhuc12) <- tolower(names(MAhuc12))
names(CThuc12) <- tolower(names(CThuc12))
names(NHhuc12) <- tolower(names(NHhuc12))
names(VThuc12) <- tolower(names(VThuc12))
names(RIhuc12) <- tolower(names(RIhuc12))

#rbind HUC12
dat <- bind_rows(MAhuc12, VThuc12, NHhuc12, CThuc12, RIhuc12)

rm(MAhuc12, CThuc12, VThuc12, NHhuc12, RIhuc12)
length(unique(dat$huc12)) #1403 

#there are duplicate entries becaseu some watersheds cross state lines
dat12 <- dat[!duplicated(dat$huc12),]
unique(dat12$name)




#load NHDflowline data
MAline <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/MA/Shape/NHDflowline.shp")
MAline2 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/MA/Shape/NHDflowline2.shp")
VTline <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/VT/Shape/NHDflowline.shp")
NHline <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/NH/Shape/NHDflowline.shp")
CTline <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/CT/Shape/NHDflowline.shp")
RIline <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/RI/Shape/NHDflowline.shp")

#need to make the names all lowercase
names(MAline) <- tolower(names(MAline))
names(MAline2) <- tolower(names(MAline2))
names(VTline) <- tolower(names(VTline))
names(NHline) <- tolower(names(NHline))
names(CTline) <- tolower(names(CTline))
names(RIline) <- tolower(names(RIline))

#rbind flowlines
flowline <- bind_rows(MAline, MAline2, CTline, RIline, VTline, NHline)

rm(MAline, MAline2, CTline, RIline, VTline, NHline)
length(unique(flowline$reachcode)) #245577

#there are duplicate entries because some watersheds cross state lines
flowline <- flowline[!duplicated(flowline$reachcode),]
flowline <- flowline %>% st_zm()







#save the merged huc files
st_write(dat8, dsn = "C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/mergedfiles/huc8.shp")
st_write(dat10, dsn = "C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/mergedfiles/huc10.shp")
st_write(dat12, dsn = "C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/mergedfiles/huc12.shp")
st_write(flowline, dsn = "C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/mergedfiles/flowline.shp")

```



here, we want to attribute each spp code with a HUC 8, 10, 12, flow line ID

```{r}
#load the merged HUC boundary files
dat8 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/mergedfiles/huc8.shp")
dat10 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/mergedfiles/huc10.shp")
dat12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/mergedfiles/huc12.shp")
flowline <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NDH/mergedfiles/flowline.shp")


#load fish data
load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_event.RData")




#make the fish data frame an sf object so it can be plotted spatially
event <- event %>% 
  filter(!is.na(latitude))
shp <- st_as_sf(x = event,                         
                coords = c("longitude", "latitude"),
                crs = st_crs(dat8))



#add the HUCs and flowline name to the fish event dataframe
shp <- st_join(shp, left = TRUE, dat8["name"])
names(shp)[10] <- "huc8name"

shp <- st_join(shp, left = TRUE, dat10["name"])
names(shp)[11] <- "huc10name"

shp <- st_join(shp, left = TRUE, dat12["name"])
names(shp)[12] <- "huc12name"

# need tofigure this out with nearest lines
#shp <- st_join(shp, left = TRUE, flowline["name"])



#plot fish distriubtion on a map to view annually
#first need to make it a spatial file and then load in watershed data, and stream data, and state data




st_write(shp, dsn = "C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/sppdata/all_fish_event.shp")


```


```{r}
shp <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/sppdata/all_fish_event.shp")

#join event data to each huc shapefile
shp <- data.frame(shp) #need to make this a dataframe in order to join with an sf object


load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_count.RData")

dat <- fish_count %>% 
  left_join( shp, by = "UID") %>% 
  filter(year >= 2015 & year <=2020)

dat1 <- dat %>% 
  filter(huc8name == "Chicopee") %>% 
  group_by(common_name) %>% 
  summarise(count = sum(count))
unique(dat1$common_name)

dat2 <- dat %>% 
  filter(huc10name == "Quaboag River")%>% 
  group_by(common_name) %>% 
  summarise(count = sum(count))
unique(dat2$common_name)


dat3 <- dat %>% 
  filter(huc12name == "Headwaters East Branch Swift River")%>% 
  group_by(common_name) %>% 
  summarise(count = sum(count))
unique(dat3$common_name)
```



here we want to plot:
the species richness by HUC size
number of surveys that have occured at each HUC size



```{r}
shp <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/sppdata/all_fish_event.shp")

#join event data to each huc shapefile
shp <- data.frame(shp) #need to make this a dataframe in order to join with an sf object


load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_count.RData")


dat <- fish_count %>% 
  left_join( shp, by = "UID") %>% 
  select(year, common_name, huc8name, huc10name, huc12name) %>% 
  unique() %>%  #need to do unique because some surveys has multiple runs so the same spp could be recorded for the same UID
  group_by(year, huc10name) %>% 
  mutate(huc10 = n_distinct(common_name)) %>% 
  ungroup() %>% 
  group_by(year, huc12name) %>% 
  mutate(huc12 = n_distinct(common_name)) %>% 
  ungroup() %>% 
  group_by(year, huc8name) %>% 
  mutate(huc8 = n_distinct(common_name)) %>% 
  ungroup() %>% 
  pivot_longer(cols = 6:8, names_to = "area", values_to = "richness")
dat$area <- factor(dat$area, levels = c("huc12", "huc10", "huc8"))


ggplot(data = dat, aes(x = area, y = richness))+
  geom_boxplot()+
  theme(axis.line = element_line(colour = "black"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "grey90"))+
  labs(x = "Area", y = "Richness")

ggsave(
    filename = "tmpfigures/richnessbyhuc.png",
    plot = last_plot(),
    width = 13,
    height = 8,
    units = "cm",
    dpi = 300
  )




#HUC8
names(dat8)[13] <- "huc8name"

test <- dat8 %>% #join the fish event dataframe to the huc 8 object
  left_join(shp, by = "huc8name") %>%
  data.frame() %>% #make a dataframe so it runs faster
  group_by(huc8name) %>%
  summarise(count = n()) #count the number of surveys per huc8 over the entire timeline of the data

dat8 <- dat8 %>% #join the counts with the spatial object so we can plot it.
  left_join(test, by = "huc8name") %>% 
  select(huc8name, count) %>% 
  unique()

ggplot(data = dat8)+ #plot of the number of surveys at each HUC8
  geom_sf(aes(fill = count, color = count))
  
ggsave(
    filename = "tmpfigures/huc8.png",
    plot = last_plot(),
    width = 13,
    height = 15,
    units = "cm",
    dpi = 300
  )

quantile(dat8$count, c(.25, .5, .75, .99))



#plot of the number of years of surveys at each HUC8
test <- dat8 %>% #join the fish event dataframe to the huc 8 object
  left_join(shp, by = "huc8name") %>%
  data.frame() %>% #make a dataframe so it runs faster
  select(huc8name, year) %>% 
  unique() %>% 
  group_by(huc8name) %>%
  summarise(count = n()) #count the number of years surveys were done at each watershed

dat8 <- dat8 %>% #join the counts with the spatial object so we can plot it.
  left_join(test, by = "huc8name") %>% 
  select(huc8name, count) %>% 
  unique()
ggplot(data = dat8)+ #plot of the number of years of surveys at each HUC8
  geom_sf(aes(fill = count, color = count))
  





#HUC10
names(dat10)[13] <- "huc10name"

test <- dat10 %>% #join the fish event dataframe to the huc 8 object
  left_join(shp, by = "huc10name") %>%
  data.frame() %>% #make a dataframe so it runs faster
  group_by(huc10name) %>%
  summarise(count = n()) #count the number of surveys per huc8 over the entire timeline of the data

dat10 <- dat10 %>% #join the counts with the spatial object so we can plot it.
  left_join(test, by = "huc10name") %>% 
  select(huc10name, count) %>% 
  unique()

ggplot(data = dat10)+ #plot of the number of surveys at each HUC8
  geom_sf(aes(fill = count, color = count))

ggsave(
    filename = "tmpfigures/huc10.png",
    plot = last_plot(),
    width = 13,
    height = 15,
    units = "cm",
    dpi = 300
  )

quantile(dat10$count, c(.25, .5, .75, .99))

#HUC12
names(dat12)[13] <- "huc12name"

test <- dat12 %>% #join the fish event dataframe to the huc 8 object
  left_join(shp, by = "huc12name") %>%
  data.frame() %>% #make a dataframe so it runs faster
  group_by(huc12name) %>%
  summarise(count = n()) #count the number of surveys per huc8 over the entire timeline of the data

dat12 <- dat12 %>% #join the counts with the spatial object so we can plot it.
  left_join(test, by = "huc12name") %>% 
  select(huc12name, count) %>% 
  unique()

ggplot(data = dat12)+ #plot of the number of surveys at each HUC8
  geom_sf(aes(fill = count, color = count))

ggsave(
    filename = "tmpfigures/huc12.png",
    plot = last_plot(),
    width = 13,
    height = 15,
    units = "cm",
    dpi = 300
  )

quantile(dat12$count, c(.25, .5, .75, .99))



table(test$common_name)

for( i in 1:length(unique(test$common_name))){
  
  ggplot()+
    geom_sf(data = MAhuc8, color = "green", fill = NA)+
    geom_sf(data = test[test$common_name == unique(test$common_name)[i], ], color = "red")+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank())+
    ggtitle(label = paste(unique(test$common_name)[i], (unique(test$scientific_name)[i]), sep = ", "))
  
  ggsave(
    filename = paste("MA fish plots/", unique(test$common_name)[i], ".png", sep = ""),
    plot = last_plot(),
    width = 9,
    height = 7,
    units = "cm",
    dpi = 300
  )
  
  print(i)
  
}


```
