---
title: "Analysis"
author: "Jenny Rogers"
date: "5/6/2022"
output: html_document
---


Install libraries

```{r}

library(tidyverse)
library(sf)
library(lwgeom)
```


load data:
flowline
huc12
huc10
huc8
all_fish_event

```{r}
flowline <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/NHDflowline/NHDflowline_NE.shp")
huc8 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU8/WBDHU8_NE.shp")
huc10 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU10/WBDHU10_NE.shp")
huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")

fish_event <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/sppdata/all_fish_event.shp")
```



spatially join the huc and the flowline data to the all_fish_event  

```{r}

#prepare the huc files - select the attributes we want and remove the duplicated rows
#21 survey points are lost from falling outside the watershed boundaries that were clipped to the state outlines.

huc8 <- huc8 %>% 
  select(tnmid, areasqkm, name) %>% 
  rename(huc8_areasqkm = areasqkm, huc8_name= name, huc8_tnmid = tnmid)
huc8 <- huc8[!duplicated(huc8$tnmid),]#Remove the duplicated TNMID fields



huc10 <- huc10 %>% 
  select(TNMID, AreaSqKm , Name)%>% 
  rename(huc10_areasqkm = AreaSqKm, huc10_name= Name, huc10_tnmid = TNMID)
huc10 <- huc10[!duplicated(huc10$TNMID),]#Remove the duplicated TNMID fields



huc12 <- huc12 %>% 
  select(tnmid, areasqkm , name) %>% 
  rename(huc12_areasqkm = areasqkm, huc12_name= name, huc12_tnmid = tnmid)
huc12 <- huc12[!duplicated(huc12$tnmid),]#Remove the duplicated TNMID fields


flowline <- flowline %>% st_zm() %>% 
  select(-c(VPUID, FDate, Resolution, FType, FCode, MainPath, InNetwork, Visibility, SlopeLenKm, PathLength))



#join the huc files to the fish_event file using a spatial join
#each fish survey event will be associated with a huc8, huc10, and huc12 TNMID
sf_use_s2(FALSE)
df <- st_join(fish_event, huc8, left = FALSE)
df <- st_join(df, huc10, left = FALSE)
df <- st_join(df, huc12, left = FALSE)



#for the flowline, we want to join the point to the nearest flowline, so we need a different approach.
nearest <- st_nearest_feature(fish_event, flowline) #returns a vector of the index of the flowline that is nearest to each point
dist <- st_distance(fish_event, flowline[nearest,], by_element = TRUE) #get dist from each point to the flowline
dist <- as.numeric(dist)
hist(dist)
fish_event_flowline_join <- cbind(fish_event, st_drop_geometry(flowline)[nearest,]) #join flowline to the fish event based on the index value of the 
fish_event_flowline_join$dist <- dist #add the distence in meters
fish_event_flowline_join <- fish_event_flowline_join %>%  #just keep lotic sites and remove all lotic sites > than 500m from the stream. I looked on GIS at some examples and 500m is about the point where is becomes obvious which site goes with which stream.
  filter(waterbody == "lotic",
         dist <500) %>% 
  select(UID, state, date, year, month, waterbody, project, source, NHDPlusID, StreamOrde, StreamLeve, AreaSqKm, TotalDASqK, Slope, dist) %>% 
  rename(event_to_flowln_dist_m = dist)
```

In this chunk of code, we prepare the presence /absence file.  Only fish presence is currently reported. We want to select only the surveys that were total pick up, and then assign a zero to all fish not observed in each survey. To clarify, all surveys will have a row for every fish and an associated 0 (negative observation) for the fish that were not observed during this survey.

```{r}

#load the fish presence absence data
load('C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_presence.RData')

fish <- unique(fish_presence$common_name)

#remove fish that are marine, estuarine, or that have only been observed in 1 location.
remove <- c("summer flounder","butterfish","king mackerel","northern kingfish",
            "bluefish","winter flounder","striped searobin","northern pipefish",
            "bay anchovy","spot","spotfin mojarra","atlantic tomcod","rainwater killifish",
            "spotfin killifish","crevalle jack","weakfish","northern searobin",
            "sheepshead minnow","atlantic needlefish","naked goby","inland silverside",
            "atlantic menhaden","three-spined stickleback","hogchoker","striped killifish",
            "atlantic silverside","striped bass","mummichog","fourspine stickleback",
            "white perch","shortnose sturgeon","white crappie","lake whitefish","redear sunfish",
            "brook silverside","northern brook lamprey","greater redhorse","guppy")


fish_presence <- fish_presence %>% 
  filter(!common_name %in% remove,
         !is.na(common_name)) 
  

uid <- unique(fish_presence$UID)
fish <- unique(fish_presence$common_name)

#make a dataframe that has each UID repeated for each fish spp
absence <- data.frame('UID' = rep(uid, each = 87))
absence$common_name <- rep(fish, times = 23348)
#keep only rows in the absence list that are no in the fish_presence list
keep <- anti_join(absence, fish_presence, by = c("UID", "common_name"))
#add additional scientific names to the new absence data frame
scientific_name <- fish_presence %>% 
  select(common_name, scientific_name, genus) %>% 
  unique()
keep <- left_join(keep, scientific_name, by = "common_name")

#add presence absence information
keep$occurrence <- 0
fish_presence$occurrence <- 1

#bind the presence and absence columns together
fish_occurrence <- rbind(fish_presence, keep)
fish_occurrence <- fish_occurrence %>% 
  arrange(UID)

rm(absence, fish_presence, keep, scientific_name)
rm(fish, uid, remove)


#read in method data and keep only the electrofish and total pick up - this will give a list of UIDs that we keep in the presence absence data (won't want to assume all other spp are absent if the survey was a targeted survey, for example). number of passes doenst really matter here because its just if the fish was there, not counts or abundance
load('C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata/all_fish_method.RData')
table(method$gear)
table(method$goal)
method <- method %>% 
  filter(goal == "total pick-up",
         gear %in% c("efish_backpack", "efish_barge", "efish_boat", "efish_misc"))

uid <- unique(method$UID)

#filter fish_occurrence data to only include UIDs in the filtered method
fish_occurrence <- fish_occurrence %>% 
  filter(UID %in% uid)

#fish_occurrence df is the file to use for the presence absence analysis.

```


In this chunk of code, we join the fish presence data to the fish_event_flowline_join, and the df dataframe so that each fish observation is associated with the hydrologic information.