---
title: "musseldataexploration"
author: "Jenny Rogers"
date: '2022-06-14'
output: html_document
---

```{r}
 

library(sf)
library(tidyverse)

```



load datafiles

Load datasets
```{r}
#watershed data
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/mussel_event_huc_join.RData")
load(file = "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/mussel_event_flowline_join.RData")


#mussel data
load(file =  "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/mussel_occurrence.RData")
load(file =  "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/mussel_count_with_zeros.RData")
load(file =  "C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata_mussel/all_mussel_length.RData")

#load new england state data
states <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/newenglandshape/NEWENGLAND_POLY.shp")

#watershed data
huc8 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU8/WBDHU8_NE.shp")
huc10 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU10/WBDHU10_NE.shp")
huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
flowline <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/NHDflowline/NHDflowline_NE.shp")


```



calcuate the number of repeat surveys at a particular NHD flowline, and each HUC size
```{r}
#add the year to each event dataset
mussel_event_flowline_join <- mussel_event_flowline_join %>% 
  mutate(year = year(date))
mussel_event_huc_join <- mussel_event_huc_join %>% 
  mutate(year = year(date))

dat <- mussel_event_flowline_join %>%
  data.frame() %>% 
  group_by(NHDPlusID) %>% 
  summarise(count = length(unique(year)))

length(unique(mussel_event_flowline_join$NHDPlusID))
hist(dat$count[dat$count>1])

nrow(dat[dat$count>10,])
nrow(dat[dat$count>5,])
nrow(dat[dat$count>1,])

dat <- mussel_event_huc_join %>% 
  data.frame() %>% 
  group_by(huc12_name) %>% 
  summarise(count = length(unique(year)))

length(unique(mussel_event_huc_join$huc12_tnmid))
hist(dat$count[dat$count>1])

nrow(dat[dat$count>10,])
nrow(dat[dat$count>5,])
nrow(dat[dat$count>1,])



# plot the number of survyes that have occurred in each HUC 12, 10, 8

#HUC12
dat <- mussel_event_huc_join %>% 
  data.frame() %>% 
  group_by(huc12_name) %>% 
  summarise(surveynum = length(unique(UID))) %>% 
  mutate(count = ifelse(surveynum %in% c(1), "1",
                        ifelse(surveynum %in% c(2:5), "2-5", 
                               ifelse(surveynum %in% c(6:15), "6-15", "16-20"))))

dat$count <- factor(dat$count, levels = c("1", "2-5", "6-15", "16-20"))


dat <- left_join(huc12, dat, by = c("name" = "huc12_name"))

 ggplot()+
    geom_sf(data = dat[!is.na(dat$surveynum),], 
            aes(fill = count), color = "black", lwd = .15)+
    geom_sf(data = states, fill = NA, lwd = 0.5, color = "black")+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank())+
    ggtitle(label = "HUC12 surveys")+
    labs(fill = "Survey Count")
 
 
 ggsave(
    filename = "C:/Users/jenrogers/Documents/git/FreshwaterBio/FreshwaterBio/tmpfigures/surveynum_mussel_huc12.png",
    plot = last_plot(),
    width = 14,
    height = 9,
    units = "cm",
    dpi = 300
  )
 
 
 
 #HUC10
dat <- mussel_event_huc_join %>% 
  data.frame() %>% 
  group_by(huc10_name) %>% 
  summarise(surveynum = length(unique(UID))) %>% 
  mutate(count = ifelse(surveynum %in% c(1:3), "1-3",
                        ifelse(surveynum %in% c(4:10), "4-10",
                               ifelse(surveynum %in% c(11:50), "11-50", ">50"))))

dat$count <- factor(dat$count, levels = c("1-3", "4-10", "11-50", ">50"))


dat <- left_join(huc10, dat, by = c("Name" = "huc10_name"))

 ggplot()+
    geom_sf(data = dat[!is.na(dat$surveynum),], 
            aes(fill = count), color = "black", lwd = 0.15)+
    geom_sf(data = states, fill = NA, lwd = 0.5, color = "black")+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank())+
    ggtitle(label = "HUC10 surveys")+
    labs(fill = "Survey Count")
 
 ggsave(
    filename = "C:/Users/Jenrogers/Documents/git/FreshwaterBio/FreshwaterBio/tmpfigures/surveynum_mussel_huc10.png",
    plot = last_plot(),
    width = 14,
    height = 9,
    units = "cm",
    dpi = 300
  )
 
 
 
 
  #HUC8
dat <- mussel_event_huc_join %>% 
  data.frame() %>% 
  group_by(huc8_name) %>% 
  summarise(surveynum = length(unique(UID))) %>% 
  mutate(count = ifelse(surveynum %in% c(1:50), "1-50",
                        ifelse(surveynum %in% c(51:100), "51:100",
                               ifelse(surveynum %in% c(101:150), "101:150", ">150"))))

dat$count <- factor(dat$count, levels = c("1-50", "51:100", "101:150", ">150"))


dat <- left_join(huc8, dat, by = c("name" = "huc8_name"))

 ggplot()+
    geom_sf(data = dat[!is.na(dat$surveynum),], 
            aes(fill = count), color = "black", lwd = 0.15)+
    geom_sf(data = states, fill = NA, lwd = 0.5, color = "black")+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank())+
    ggtitle(label = "HUC8 surveys")+
    labs(fill = "Survey Count")
 
 ggsave(
    filename = "C:/Users/jenrogers/Documents/git/FreshwaterBio/FreshwaterBio/tmpfigures/surveynum_mussel_huc8.png",
    plot = last_plot(),
    width = 14,
    height = 9,
    units = "cm",
    dpi = 300
  )
 
```



Plot the fish distributions 
These plots include all the surveys, included the lentic surveys and the ones > 500 m
```{r}

#load new englad state data
states <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/newenglandshape/NEWENGLAND_POLY.shp")
rivers <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/NHDflowline_RiiversOnly_NE.shp")

shp <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/sppdata/all_mussel_occurrence.shp")

shp <- shp %>% 
  filter(lv_ccrr == 1)



#plot fish distriubtion on a map to view annually
#first need to make it a spatial file and then load in watershed data, and stream data, and state data

table(shp$cmmn_nm)

for( i in 1:length(unique(shp$cmmn_nm))){
  
  ggplot()+
    geom_sf(data = states)+
    geom_sf(data = huc8, fill = NA, lwd = 0.25)+
    geom_sf(data = shp[shp$cmmn_nm == unique(shp$cmmn_nm)[i], ], color = "red", size = .5)+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank())+
    ggtitle(label = paste(unique(shp$cmmn_nm)[i], (unique(shp$scntfc_)[i]), sep = ", "))
  
  ggsave(
    filename = paste("tmpfigures/muss_dist_plots/", unique(shp$cmmn_nm)[i], ".png", sep = ""),
    plot = last_plot(),
    width = 14,
    height = 9,
    units = "cm",
    dpi = 300
  )
  
  print(i)
  
}


VT <- states %>% 
  filter(NAME == "VERMONT")


```
number of unique survey events

```{r}
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata_mussel/all_mussel_event.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata_mussel/all_mussel_occurrence.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata_mussel/all_mussel_count.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata_mussel/all_mussel_method.RData")
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata_mussel/all_mussel_length.RData")

tmp <- all_mussel_event %>% 
  group_by(state) %>% 
  summarise(count = n())

unique(all_mussel_method$survey_method)

#number of sureys with occurrence vs counts
length(unique(all_mussel_count$UID))
#1544
length(unique(all_mussel_occurrence$UID))
#3789



```


surveys by year and state
```{r}
load("C:/Users/jenrogers/Documents/necascFreshwaterBio/spp_data/tidydata_mussel/all_mussel_event.RData")

#surveys by year and state
all_mussel_event$year <- year(all_mussel_event$date)
table(all_mussel_event$year)
all_mussel_event$year[all_mussel_event$year == 1 & !is.na(all_mussel_event$year)] <- sample(1980:2005, length(all_mussel_event$year[all_mussel_event$year == 1]), replace = T)


ggplot(data = all_mussel_event[all_mussel_event$year>1975 & all_mussel_event$state !="NY",], mapping = aes(x = year))+
  geom_bar(aes(fill = state))+
  theme(axis.line = element_line(colour = "black"),
        panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(color = "grey90"))+
  labs(x = "Year", y = "Count", fill = "State")
ggsave(
    filename = "tmpfigures/musselsurveybyyear.png",
    plot = last_plot(),
    width = 15,
    height = 10,
    units = "cm",
    dpi = 300
  )
  



```


plots of presence and absence
plots of counts
These plots only include the surveys of lotic sites that are <500m from a stream
```{r}
head(mussel_occurrence)

#load mussel data
shp <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/sppdata/all_mussel_occurrence.shp")
#get the UIDs in the mussel_event_huc_join because these have been editted to remove lentic and to remove sites >500m from stream



#remove NA names
shp <- shp %>% 
  filter(UID %in% mussel_event_huc_join$UID) %>% #remove the lentic sites
  select(UID) %>% 
  unique()



####Presence absence:



pres <- right_join(shp, mussel_occurrence, by = "UID") %>% 
  filter(!is.na(live_occurrence))


table(pres$common_name)

for( i in 1:length(unique(pres$common_name))){
  
  ggplot()+
    geom_sf(data = states)+
    geom_sf(data = pres[pres$common_name== unique(pres$common_name)[i], ], 
            aes(color = as.factor(live_occurrence)), size = 1)+
    geom_sf(data = huc8, fill = NA, color = "black", lwd = 0.15)+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank())+
    ggtitle(label = unique(pres$common_name)[i])+
    labs(color = "Occurrence")
  
  ggsave(
    filename = paste("tmpfigures/mussel_occ_with_zero/", unique(pres$common_name)[i], ".png", sep = ""),
    plot = last_plot(),
    width = 14,
    height = 9,
    units = "cm",
    dpi = 300
  )
  
  print(i)
  
}


#### proportional abundance

#mussels with counts:
temp <- mussel_count_with_zeros   %>% 
  filter(!is.na(live_count),
         live_count>0) %>%  
  group_by(UID) %>% 
  mutate(total_count = sum(live_count),
         prop_abun = live_count/total_count) 

temp2 <- mussel_count_with_zeros %>% 
  filter(live_count == 0) %>% 
  mutate(prop_abun = 0)
#need to add the total counts to temp2
temp3 <- temp %>% 
  select(UID, total_count) %>% 
  unique()

temp2 <- left_join(temp2, temp3, by = "UID") %>% 
  arrange(UID)

counts <- rbind(temp, temp2) %>% 
  arrange(UID)
counts$total_count <- ifelse(is.na(counts$total_count), 0, counts$total_count)
  

counts <- right_join(shp, counts, by = "UID") 

counts$occurrence <- as.factor(ifelse(counts$live_count>0, 1, 0))


table(counts$common_name)

for( i in 1:length(unique(counts$common_name))){
  
  ggplot()+
    geom_sf(data = states)+
    geom_sf(data = counts[counts$common_name== unique(counts$common_name)[i], ],
            pch = 21,
            aes(size = prop_abun, fill = occurrence))+
    geom_sf(data = huc8, fill = NA, color = "black", lwd = 0.15)+
  scale_size(limits = c(0,1))+
    scale_fill_manual(values =c("1" = "aquamarine", "0" = "black"))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank())+
    ggtitle(label = unique(counts$common_name)[i])+
    labs(size = "prop abun", fill = "occurrence")
  
  ggsave(
    filename = paste("tmpfigures/mussel_count_with_zero/", unique(counts$common_name)[i], ".png", sep = ""),
    plot = last_plot(),
    width = 14,
    height = 9,
    units = "cm",
    dpi = 300
  )
  
  print(i)
  
}


```





Species richness by HUC plots
```{r}

huc8 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU8/WBDHU8_NE.shp")
huc10 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU10/WBDHU10_NE.shp")
huc12 <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/NHDplus/WBDHU12/WBDHU12_NE.shp")
shp <- st_read("C:/Users/jenrogers/Documents/necascFreshwaterBio/SpatialData/sppdata/all_mussel_occurrence.shp")



huc8 <- huc8 %>% 
  select(tnmid, areasqkm, name, states) %>% 
  rename(huc8_areasqkm = areasqkm, huc8_name= name, huc8_tnmid = tnmid)
huc8 <- huc8[!duplicated(huc8$huc8_tnmid),]#Remove the duplicated TNMID fields



huc10 <- huc10 %>% 
  select(TNMID, AreaSqKm , Name, States)%>% 
  rename(huc10_areasqkm = AreaSqKm, huc10_name= Name, huc10_tnmid = TNMID)
huc10 <- huc10[!duplicated(huc10$huc10_tnmid),]#Remove the duplicated TNMID fields



huc12 <- huc12 %>% 
  select(tnmid, areasqkm , name, states) %>% 
  rename(huc12_areasqkm = areasqkm, huc12_name= name, huc12_tnmid = tnmid)
huc12 <- huc12[!duplicated(huc12$huc12_tnmid),]#Remove the duplicated TNMID fields



#join the huc files to the fish_event file using a spatial join
#each fish survey event will be associated with a huc8, huc10, and huc12 TNMID
sf_use_s2(FALSE)
mussel_event_huc_join <- st_join(shp, huc8, left = FALSE)
mussel_event_huc_join <- st_join(mussel_event_huc_join, huc10, left = FALSE)
mussel_event_huc_join <- st_join(mussel_event_huc_join, huc12, left = FALSE)

#spp richness: huc12

richness_huc12 <- mussel_event_huc_join %>% 
  filter(lv_ccrr == 1) %>%  
  data.frame() %>% 
  group_by(huc12_name) %>%
  summarise(richess_huc12 = length(unique(cmmn_nm)))
  
  

hist(richness_huc12$richess_huc12)

huc12 <- huc12 %>% 
  filter(states != "CN")

dat <- left_join(huc12, richness_huc12, by = "huc12_name")

hist(dat$richess_huc12)

dat$richess_huc12 <- as.numeric(dat$richess_huc12)

ggplot()+
  geom_sf(data = dat, lwd = 0.2,
          aes(fill = richess_huc12))+
  scale_fill_gradient2(low = "red",
  mid = "white",
  high = "blue",
  midpoint = 5)+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank())+
    labs(title = "species richness, huc12 scale",
    fill = "richness")+
  geom_sf(data = states, fill = NA, lwd = 0.5, color = "black")

 ggsave(
    filename = "tmpfigures/musselspprichness_huc12_scale.png",
    plot = last_plot(),
    width = 14,
    height = 9,
    units = "cm",
    dpi = 250
  )
 
 
 
 #spp richness: huc10

richness_huc10 <- mussel_event_huc_join %>% 
  filter(lv_ccrr == 1) %>%  
  data.frame() %>% 
  group_by(huc10_name) %>%
  summarise(richness_huc10 = length(unique(cmmn_nm)))
  
  
hist(as.numeric(richness_huc10$richness_huc10))


huc10 <- huc10 %>% 
  filter(States != "CN")

dat <- left_join(huc10, richness_huc10, by = "huc10_name")

hist(dat$richness_huc10)

dat$richness_huc10 <- as.numeric(dat$richness_huc10)

ggplot()+
  geom_sf(data = dat, 
          aes(fill = richness_huc10))+
  scale_fill_gradient2(low = "red",
  mid = "white",
  high = "blue",
  midpoint = median(dat$richness_huc10, na.rm = T))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank())+
    labs(title = "species richness, huc10 scale",
    fill = "richness")+
  geom_sf(data = states, fill = NA, lwd = .5, color = "black")

 ggsave(
    filename = "tmpfigures/musselspprichness_huc10_scale.png",
    plot = last_plot(),
    width = 14,
    height = 9,
    units = "cm",
    dpi = 250
  )
 
 
 
 
 
 #spp richness: huc8

richness_huc8 <- mussel_event_huc_join %>% 
  filter(lv_ccrr == 1) %>%  
  data.frame() %>% 
  group_by(huc8_name) %>%
  summarise(richness_huc8 = length(unique(cmmn_nm)))
  
  
hist(as.numeric(richness_huc8$richness_huc8))


huc8 <- huc8 %>% 
  filter(states != "CN")

dat <- left_join(huc8, richness_huc8, by = "huc8_name")

hist(dat$richness_huc8)

dat$richness_huc8 <- as.numeric(dat$richness_huc8)

ggplot()+
  geom_sf(data = dat, 
          aes(fill = richness_huc8))+
  scale_fill_gradient2(low = "red",
  mid = "white",
  high = "blue",
  midpoint = median(dat$richness_huc8, na.rm = T))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank())+
    labs(title = "species richness, huc8 scale",
    fill = "richness")+
  geom_sf(data = states, fill = NA, lwd = .5, color = "black")

 ggsave(
    filename = "tmpfigures/musselspprichness_huc8_scale.png",
    plot = last_plot(),
    width = 14,
    height = 9,
    units = "cm",
    dpi = 250
  )
 
 ## spp richness, beta - alpha / gamma
 
 #want to divide the huc 12 richness by the huc 8 richness
 dat <- left_join(mussel_event_huc_join, richness_huc12, by = "huc12_name")
 dat <- left_join(dat, richness_huc8, by = "huc8_name")
 dat <- dat %>% 
   data.frame() %>% 
   select(richess_huc12, richness_huc8, huc12_name) %>% 
   unique()
 dat$beta <- dat$richess_huc12/ dat$richness_huc8

 huc12 <- huc12 %>% 
  filter(states != "CN")
 
 dat <- left_join(huc12, dat, by = "huc12_name")

hist(dat$beta)

dat$beta <- as.numeric(dat$beta)
#need to figure out why a handful of HUC12s have a greater spp richness than the huc8.. not possible
ggplot()+
  geom_sf(data = dat[dat$beta<=1,], lwd = 0.2,
          aes(fill = beta))+
  scale_fill_gradient2(low = "red",
  mid = "yellow",
  high = "green",
  midpoint = 0.5)+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank())+
    labs(title = "beta diversity = alpha/gamma",
    fill = "diversity")+
  geom_sf(data = states, fill = NA, lwd = 0.25, color = "black")+
  geom_sf(data = huc8, fill = NA, lwd = 0.25, color = "black")

 ggsave(
    filename = "tmpfigures/mussel_spprichness_beta.png",
    plot = last_plot(),
    width = 14,
    height = 9,
    units = "cm",
    dpi = 300
  )

 
 

```



Shannon Wiener and Simpsons Index
```{r}


#need to get the number of surveys in each HUC so we can divide abundance by the number of surveys
surveycounts <- left_join(mussel_count_with_zeros, 
                          mussel_event_huc_join, 
                          by = "UID") %>% 
   select(UID, huc10_name) %>% 
   unique() %>% 
   group_by(huc10_name) %>% 
   summarise(suveynum = n())



#### Shannon Wiener



shannon1 <- mussel_count_with_zeros %>% 
  filter(live_count>0,
         !is.na(live_count),
         !is.na(common_name))

#count of each spp by huc
shannon <- left_join(mussel_event_huc_join, shannon1, by = "UID") %>%
  filter(!is.na(live_count),
         !is.na(common_name)) %>% 
  group_by(huc10_name, common_name) %>% 
  summarise(n = sum(live_count)) %>% 
  data.frame()

shannon <- left_join(shannon, surveycounts, by = "huc10_name") %>% 
  mutate(n1 = n/suveynum)

#Total sum of all counts of all spp
shannon2 <- left_join(mussel_event_huc_join, shannon1, by = "UID") %>%
  filter(!is.na(live_count),
         !is.na(common_name)) %>% 
  group_by(huc10_name)%>% 
  summarise(N = sum(live_count)) %>% 
  data.frame()

shannon2 <- left_join(shannon2, surveycounts, by = "huc10_name") %>% 
  mutate(N1 = N/suveynum)

shannon3 <- left_join(shannon, shannon2, by = "huc10_name") %>% 
  mutate(tmp = n1/N1,
         tmp2 = log(tmp),
         tmp3 = tmp * tmp2) %>% 
  group_by(huc10_name) %>% 
  summarise(shannonwiener = -(sum(tmp3)))




huc10 <- huc10 %>% 
  filter(States != "CN")

dat <- left_join(huc10, shannon3, by = "huc10_name")

hist(dat$shannonwiener)

dat$shannonwiener <- as.numeric(dat$shannonwiener)

ggplot()+
  geom_sf(data = dat, lwd = 0.2,
          aes(fill = shannonwiener))+
  scale_fill_gradient2(low = "red",
  mid = "white",
  high = "blue",
  midpoint = mean(dat$shannonwiener, na.rm = T))+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank())+
    labs(title = "Shannon Wiener (H`), huc10 scale",
    fill = "H`")+
  geom_sf(data = states, fill = NA, lwd = 0.25, color = "black")+
  geom_sf(data = huc8, fill = NA, lwd = 0.25, color = "black")


 ggsave(
    filename = "tmpfigures/mussel_shannonwiener_huc10.png",
    plot = last_plot(),
    width = 14,
    height = 9,
    units = "cm",
    dpi = 300
  )
 
 
 #### Simpsons Index
 
 
 simpsons <- left_join(shannon, shannon2, by = "huc10_name") %>% 
  mutate(tmp = n*(n-1),
         tmp2 = N*(N-1),
         tmp3 = tmp / tmp2) %>% 
  group_by(huc10_name) %>% 
  summarise(simpsons = 1 - (sum(tmp3)))


 
 huc10 <- huc10 %>% 
  filter(States != "CN")

dat <- left_join(huc10, simpsons, by = "huc10_name")

hist(dat$simpsons)

 


dat$simpsons <- as.numeric(dat$simpsons)

ggplot()+
  geom_sf(data = dat, lwd = 0.2,
          aes(fill = simpsons))+
  scale_fill_gradient2(low = "red",
  mid = "white",
  high = "blue",
  midpoint = 0.5)+
    theme(panel.border = element_rect(colour = "black", fill = NA),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank())+
    labs(title = "Simpsons, huc10 scale",
    fill = "Probability")+
  geom_sf(data = states, fill = NA, lwd = 0.25, color = "black")+
  geom_sf(data = huc8, fill = NA, lwd = 0.25, color = "black")

 ggsave(
    filename = "tmpfigures/mussel_Simpsons_huc10.png",
    plot = last_plot(),
    width = 14,
    height = 9,
    units = "cm",
    dpi = 300
  )


```







Cluster spp based on co - occurrence 
```{r}
library(data.table)


########################################

#cluster based on the occurrence of other species lumping all basins and stream orders

########################################



load(file =  "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/mussel_occurrence.RData")



dat <- mussel_occurrence %>%
  select(UID, common_name, live_occurrence) %>%
  group_by(UID, common_name) %>% 
  summarise(live_occurrence = sum(live_occurrence)) %>% #in the raw data, there are 21 UIDs that say a certain spp was both 0 and 1, we will keep the 1s
  pivot_wider(names_from = common_name, values_from = live_occurrence) %>% 
  column_to_rownames(var = "UID")

test <- t(dat)
colnames(test) <- rownames(dat)

dist_mat <- dist(test, method = "binary")

hclust_avg <- hclust(dist_mat, method = "average")
hclust_avg$labels<- names(dat)

png(height = 10, width = 15, file = "tmpfigures/musselhclust.png", units = "in", res = 150, type = "cairo")

plot(hclust_avg)
dev.off()






#####################################

#cluster based on the occurrence of other species for each HUC8, lumping all stream orders within each HUC 8

#####################################



load(file =  "C:/Users/jenrogers/Documents/necascFreshwaterBio/model_datafiles/mussel_occurrence.RData")


  
tmp <- left_join(mussel_occurrence, mussel_event_huc_join, by = "UID") 

for (i in unique(mussel_event_huc_join$huc8_name)) {

dat <- tmp %>% 
  filter(huc8_name == i)
#remove spp that are never present
keep <- dat %>% 
  group_by(common_name) %>% 
  summarise(occ = sum(live_occurrence)) %>% 
  filter(occ > 0)


dat <- dat %>%
  filter(common_name %in% keep$common_name ) %>%  
  select(UID, common_name, live_occurrence) %>%
  group_by(UID, common_name) %>% 
  summarise(live_occurrence = sum(live_occurrence)) %>%  
  pivot_wider(names_from = common_name, values_from = live_occurrence) %>% 
  column_to_rownames(var = "UID")

test <- t(dat)
colnames(test) <- rownames(dat)

dist_mat <- dist(test, method = "binary")

hclust_avg <- hclust(dist_mat, method = "average")
hclust_avg$labels<- names(dat)

png(height = 10, width = 15, file = paste("tmpfigures/mussel_clusters/", i, ".png", sep = ""), units = "in", res = 150, type = "cairo")

plot(hclust_avg)
dev.off()

print(i)

}




```